% !TEX program = pdflatex
% !TEX encoding = UTF-8
% !TEX spellcheck = en_GB
% !TEX root = peter-pan.tex

\Act{The Nursery}

\begin{stagedir}
The night nursery of the Darling family,
which is the scene of our opening Act,
is at the top of a rather depressed street in Bloomsbury.
We have a right to place it where we will,
and the reason Bloomsbury is chosen is that Mr.\@ Roget once lived there.
So did we in days when his \emph{Thesaurus} was our only companion in London;
and we whom he has helped to wend our way through life
have always wanted to pay him a little compliment.
The Darlings therefore lived in Bloomsbury.

It is a corner house whose top window, the important one,
looks upon a leafy square from which Peter used to fly up to it,
to the delight of three children and no doubt the irritation of passers-by.
The street is still there, though the steaming sausage shop has gone;
and apparently the same cards perch now as then over the doors,
inviting homeless ones to come and stay with the hospitable inhabitants.
Since the days of the Darlings, however, a lick of paint has been applied;
and our corner house in particular, which has swallowed its neighbour,
blooms with awful freshness as if the colours had been discharged upon it through a hose.
Its card now says 'No children'
meaning maybe that the goings-on of Wendy and her brothers have given the house a bad name.
As for ourselves, we have not been in it since we went back to reclaim our old \emph{Thesaurus}.

That is what we call the Darling house,
but you may dump it down anywhere you like, and if you think it was your house you are very probably right.
It wanders about London looking for anybody in need of it, like the little house in the Never Land.

The blind
(which is what Peter would have called the theatre curtain if he had ever seen one)
rises on that top room,
a shabby little room if Mrs.\@ Darling had not made it the hub of creation by her certainty that such it was,
and adorned it to match with a loving heart and all the scrapings of her purse.
The door on the right leads into the day nursery,
which she has no right to have, but she made it herself with nails in her mouth and a paste-pot in her hand.
This is the door the children will come in by.
There are three beds and (rather oddly) a large dog-kennel;
two of these beds, with the kennel, being on the left and the other on the right.
The coverlets of the beds (if visitors are expected) are made out of Mrs.\@ Darling's wedding-gown,
which was such a grand affair that it still keeps them pinched.
Over each bed is a china house, the size of a linnet's nest, containing a night-light.
The fire, which is on our right, is burning as discreetly as if it were in custody,
which in a sense it is, for supporting the mantelshelf are two wooden soldiers, home-made,
begun by Mr.\@ Darling, finished by Mrs.\@ Darling, repainted (unfortunately) by John Darling.
On the fire-guard hang incomplete parts of children's night attire.
The door the parents will come in by is on the left.
At the back is the bathroom door, with a cuckoo clock over it;
and in the centre is the window, which is at present ever so staid and respectable,
but half an hour hence (namely at 6.30 p.m.) will be able to tell a very strange tale to the police.

The only occupant of the room at present is Nana the nurse,
reclining, not as you might expect on the one soft chair, but on the floor.
She is a Newfoundland dog, and though this may shock the grandiose,
the not exactly affluent will make allowances.
The Darlings could not afford to have a nurse,
they could not afford indeed to have children;
and now you are beginning to understand how they did it.
Of course Nana has been trained by Mrs.\@ Darling,
but like all treasures she was born to it.
In this play we shall see her chiefly inside the house,
but she was just as exemplary outside,
escorting the two elders to school with an umbrella in her mouth, for instance,
and butting them back into line if they strayed.

The cuckoo clock strikes six, and Nana springs into life.
This first moment in the play is tremendously important,
for if the actor playing Nana does not spring properly we are undone.
She will probably be played by a boy, if one clever enough can be found,
and must never be on two legs except on those rare occasions when an ordinary nurse would be on four.
This Nana must go about all her duties in a most ordinary manner,
so that you know in your bones that she performs them just so every evening at six;
naturalness must be her passion;
indeed, it should be the aim of every one in the play, for which she is now setting the pace.
All the characters, whether grown-ups or babes,
must wear a child's outlook on life as their only important adornment.
If they cannot help being funny they are begged to go away.
A good motto for all would be 'The little less, and how much it is.'

Nana, making much use of her mouth, 'turns down' the beds,
and carries the various articles on the fire-guard across to them.
Then pushing the bathroom door open,
she is seen at work on the taps preparing Michael's bath;
after which she enters from the day nursery with the youngest of the family on her back.
\end{stagedir}

\begin{drama}

\michaelspeaks \direct{obstreperous}¤
I won't go to bed, I won't, I won't.
Nana, it isn't six o'clock yet.
Two minutes more, please, one minute more?
Nana, I won't be bathed, I tell you I will not be bathed.

\end{drama}

\endinput

<p><i>(Here the bathroom door closes on them, and</i> MRS. DARLING, <i>who has perhaps heard his cry, enters the nursery. She is the loveliest lady in Bloomsbury, with a sweet mocking mouth, and as she is going out to dinner tonight she is already wearing her evening gown because she knows her children like to see her in it. It is a delicious confection made by herself out of nothing and other people's mistakes. She does not often go out to dinner, preferring when the children are in bed to sit beside them tidying up their minds, just as if they were drawers. If</i> WENDY <i>and the boys could keep awake they might see her repacking into their proper places the many articles of the mind that have strayed during the day, lingering humorously over some of their contents, wondering where on earth they picked this thing up, making discoveries sweet and not so sweet, pressing this to her cheek and hurriedly stowing that out of sight. When they wake in the morning the naughtinesses with which they went to bed are not, alas, blown away, but they are placed at the bottom of the drawer; and on the top, beautifully aired, are their prettier thoughts ready forthe new day.</i></p>

<p><i>As she enters the room she is startled to see a strange little face outside the window and a hand groping as if it wanted to come in.)</i></p>

<p>MRS. DARLlNG. Who are you?
\direct{The unknown disappears; she hurries to the window.}
No one there. And yet I feel sure I saw a face. My children!
\direct{She throws open the bathroom door and</i> MICHAEL'S <i>head appears gaily over the bath. He splashes; she throws kisses to him and closes the door.'Wendy, John,' she cries, and gets reassuring answers from the day nursery. She sits down, relieved, on</i> WENDY'S <i>bed; and</i> WENDY <i>and</i> JOHN <i>come in, looking their smallest size, as children tend to do to a mother suddenly in fear for them.}
</p>

<p>JOHN \direct{histrionically}¤
We are doing an act; we are playing at being you and father.
\direct{He imitates the only father who has come under his special notice.}
A little less noise there.</p>

<p>WENDY. Now let us pretend we have a baby.</p>

<p>JOHN \direct{good-naturedly}¤
I am happy to inform you, Mrs.Darling, that you are now a mother. (WENDY <i>gives way to ecstasy.)</i> You have missed the chief thing; you haven't asked, 'boy or girl?'</p>

<p>WENDY. I am so glad to have one at all, I don't care which it is.</p>

<p>JOHN \direct{crushingly}¤
That is just the difference between gentlemen and ladies. Now you tell me.</p>

<p>WENDY. I am happy to acquaint you, Mr. Darling, you are now a father.</p>

<p>JOHN. Boy or girl?</p>

<p>WENDY \direct{presenting herself}¤
Girl.</p>

<p>JOHN. Tuts.</p>

<p>WENDY. You horrid.</p>

<p>JOHN. Go on.</p>

<p>WENDY. I am happy to acquaint you, Mr. Darling, you are again a father.</p>

<p>JOHN. Boy or girl?</p>

<p>WENDY. Boy. (JOHN <i>beams.)</i> Mummy, it's hateful of him.</p>

<blockquote>(MICHAEL <i>emerges from the bathroom in</i> JOHN'S <i>old pyjamas and giving his face a last wipe with the towel.)</i></blockquote>

<p>MICHAEL \direct{expanding}¤
Now, John, have me.</p>

<p>JOHN. We don't want any more.</p>

<p>MICHAEL \direct{contracting}¤
Am I not to be born at all?</p>

<p>JOHN. Two is enough.</p>

<p>MICHAEL \direct{wheedling}¤
Come, John; boy, John.
\direct{Appalled.}
Nobody wants me!</p>

<p>MRS. DARLING. I do.</p>

<p>MICHAEL \direct{with a glimmer of hope}¤
Boy or girl?</p>

<p>MRS. DARLING \direct{with one of those happy thoughts of hers}¤
Boy.</p>

<blockquote>(<i>Triumph of</i> MICHAEL; <i>discomfiture of</i> JOHN. MR.DARLING <i>arrives, in no mood unfortunately to gloat over this domestic scene. He is really a good man as breadwinners go, and it is hard luck for him to be propelled into the room now, when if we had brought him in a few minutes earlier or later he might have made a fairer impression. In the city where he sits on a stool all day, as fixed as a postage stamp, he is so like all the others on stools that you recognise him not by his face but by his stool, but at home the way to gratify him is to say that he has a distinct personality. He is very conscientious, and in the days when</i> MRS. DARLING <i>gave up keeping the house books correctly and drew pictures instead (which he called her guesses), he did all the totting up for her, holding her hand while he calculated whether they could have Wendy or not, and coming down on the right side. It is with regret, therefore, that we introduce him as a tornado, rushing into the nursery in evening dress, but without his coat, and brandishing in his hand a recalcitrant white tie.)</i></blockquote>

<p>MR. DARLING \direct{implying that he has searched for her everywhere and that the nursery is a strange place in which to find her}¤
Oh, here you are, Mary.</p>

<p>MRS. DARLING \direct{knowing at once what is the matter}¤
What is the matter, George dear?</p>

<p>MR. DARLING \direct{as if the word were monstrous}¤
Matter! This tie, it will not tie.
\direct{He waxes sarcastic.}
Not round my neck. Round the bed-post, oh yes; twenty times have I made it up round the bed-post, but round my neck, oh dear no; begs to be excused.</p>

<p>MICHAEL \direct{in a joyous transport}¤
Say it again, father, say it again!</p>

<p>MR. DARLING (<i>witheringly</i>). Thank you.
\direct{Goaded by asuspiciously crooked smile on</i> MRS. DARLING'S <i>face.}
I warn you, Mary, that unless this tie is round my neck we don't goout to dinner to-night, and if I don't go out to dinner to-night I never go to the office again, and if I don't go to the office again you and I starve, and our children will be thrown into the streets.</p>

<blockquote><i>(The children blanch as they grasp the gravity of the situation.)</i></blockquote>

<p>MRS. DARLING. Let me try, dear.</p>

<blockquote><i>(In a terrible silence their progeny cluster round them. Will she succeed? Their fate depends on it. She fails</i>&mdash;<i>no, she succeeds. In another moment they are wildly gay, romping round the room on each other's shoulders. Father is even a better horse than mother.</i> MICHAEL <i>is dropped upon his bed,</i> WENDY <i>retires to prepare for hers,</i> JOHN <i>runs from</i> NANA, <i>who has reappeared with the bath towel.)</i></blockquote>

<p>JOHN \direct{rebellious}¤
I won't be bathed. You needn't think it.</p>

<p>MR. DARLING \direct{in the grand manner}¤
Go and be bathed at once, sir.</p>

<blockquote><i>(With bent head</i> JOHN <i>follows</i> NANA <i>into the bathroom.</i> MR. DARLING <i>swells.)</i></blockquote>

<p>MICHAEL \direct{as he is put between the sheets}¤
Mother, how did you get to know me?</p>

<p>MR. DARLING. A little less noise there.</p>

<p>MICHAEL \direct{growing solemn}¤
At what time was I born, mother?</p>

<p>MRS. DARLING. At two o'clock in the night-time, dearest.</p>

<p>MICHAEL. Oh, mother, I hope I didn't wake you.</p>

<p>MRS. DARLING. They are rather sweet, don't you think,George?</p>

<p>MR. DARLING \direct{doting}¤
There is not their equal on earth, and they are ours, ours!</p>

<blockquote><i>(Unfortunately</i> NANA <i>has come from the bathroom for a sponge and she collides with his trousers<sub>)</sub> the first pair</i> <i>he has ever had with braid on them.)</i></blockquote>

<p>MR. DARLING. Mary, it is too bad; just look at this; covered with hairs. Clumsy, clumsy!</p>

<blockquote>(NANA <i>goes, a drooping figure.)</i></blockquote>

<p>MRS. DARLING. Let me brush you, dear.</p>

<blockquote><i>(Once more she is successful. They are now by the fire, and</i> MICHAEL <i>is in bed doing idiotic things with a teddy bear.)</i></blockquote>

<p>MR. DARLING \direct{depressed}¤
I sometimes think, Mary, that it is a mistake to have a dog for a nurse.</p>

<p>MRS. DARLING. George, Nana is a treasure.</p>

<p>MR. DARLING. No doubt; but I have an uneasy feeling at times that she looks upon the children as puppies.</p>

<p>MRS. DARLING \direct{rather faintly}¤
Oh no, dear one, I am sure she knows they have souls.</p>

<p>MR. DARLING \direct{profoundly}¤
I wonder, I wonder.</p>

<blockquote><i>(The opportunity has come for her to tell him of something that is on her mind.)</i></blockquote>

<p>MRS. DARLING. George, we must keep Nana. I will tell you why.
\direct{Her seriousness impresses him.}
My dear, when I came into this room to-night I saw a face at the window.</p>

<p>MR. DARLING <i>(incredulous). A</i> face at the window, three floors up? Pooh!</p>

<p>MRS. DARLING. It was the face of a little boy; he was trying to get in. George, this is not the first time I have seen that boy.</p>

<p>MR. DARLING \direct{beginning to think that this may be a man's job}¤
Oho!</p>

<p>MRS. DARLING \direct{making sure that</i> MICHAEL <i>does not hear}¤
The first time was a week ago. It was Nana's night out, and I had been drowsing here by the fire when suddenly I felt a draught, as if the window were open. I looked round and I saw that boy&mdash;in the room.</p>

<p>MR. DARLING. In the room?</p>

<p>MRS. DARLING. I screamed. Just then Nana came back and she at once sprang at him. The boy leapt for the window. She pulled down the sash quickly, but was too late to catch him.</p>

<p>MR. DARLING \direct{who knows he would not have been too late}¤
I thought so!</p>

<p>MRS. DARLING. Wait. The boy escaped, but his shadow had not time to get out; down came the window and cut it clean off.</p>

<p>MR. DARLING \direct{heavily}¤
Mary, Mary, why didn't you keep that shadow?</p>

<p>MRS. DARLING \direct{scoring}¤
I did. I rolled it up, George; and here it is.</p>

<blockquote><i>(She produces it from a drawer. They unroll and examine the flimsy thing, which is not more material than a puff of smoke, and if let go would probably float into the ceiling without discolouring it. Yet it has human shape. As they nod their heads over it they present the most satisfying picture on earth, two happy parents conspiring cosily by the fire for the good of their children.)</i></blockquote>

<p>MR. DARLING. It is nobody I know, but he does look ascoundrel.</p>

<p>MRS. DARLING. I think he comes back to get his shadow,George.</p>

<p>MR. DARLING \direct{meaning that the miscreant has now a father to deal with}¤
I dare say.
\direct{He sees himself telling the story to the other stools at the office.}
There is money in this, my love. I shall take it to the British Museum to-morrow and have it priced.</p>

<blockquote><i>(The shadow is rolled up and replaced in the drawer.)</i></blockquote>

<p>MRS. DARLING \direct{like a guilty person}¤
George, I have not told you all; I am afraid to.</p>

<p>MR. DARLING \direct{who knows exactly the right moment to treat a woman as a beloved child}¤
Cowardy, cowardy custard.</p>

<p>MRS. DARLING \direct{pouting}¤
No, I 'm not.</p>

<p>MR. DARLING. Oh yes, you are.</p>

<p>MRS. DARLING. George, I 'm not.</p>

<p>MR. DARLING. Then why not tell?
\direct{Thus cleverly soothed she goes on.}</p>

<p>MRS. DARLING. The boy was not alone that first time. He was accompanied by&mdash;I don't know how to describe it; by a ball of light, not as big as my fist, but it darted about the room like a living thing.</p>

<p>MR. DARLING \direct{though open-minded}¤
That is very unusual. It escaped with the boy?</p>

<p>MRS. DARLING. Yes.
\direct{Sliding her hand into his.}
George, what can all this mean?</p>

<p>MR. DARLING \direct{ever ready}¤
What indeed!</p>

<blockquote><i>(This intimate scene is broken by the return of</i> NANA <i>with a bottle in her mouth.)</i></blockquote>

<p>MRS. DARLING \direct{at once dissembling}¤
What is that, Nana? Ah, of course; Michael, it is your medicine.</p>

<p>MICHAEL \direct{promptly}¤
Won't take it.</p>

<p>MR. DARLING \direct{recalling his youth}¤
Be a man, Michael.</p>

<p>MICHAEL. Won't.</p>

<p>MRS. DARLING \direct{weakly}¤
I'll get you a lovely chocky to take after it.
\direct{She leaves the room, though her husband calls after her.}</p>

<p>MR. DARLING. Mary, don't pamper him. When I was your age, Michael, I took medicine without a murmur. I said 'Thank you, kind parents, for giving me bottles to make me well.'</p>

<blockquote>(WENDY, <i>who has appeared in her nightgown, hears this and believes.)</i></blockquote>

<p>WENDY. That medicine you sometimes take is much nastier, isn't it, father?</p>

<p>MR. DARLING \direct{valuing her support}¤
Ever so much nastier.And as an example to you, Michael, I would take it now
\direct{thankfully}
if I hadn't lost the bottle.</p>

<p>WENDY \direct{always glad to be of service}¤
I know where itis, father. I'll fetch it.</p>

<blockquote><i>(She is gone before he can stop her. He turns for help to</i> JOHN, <i>who has come from the bathroom attired for bed.)</i></blockquote>

<p>MR. DARLING. John, it is the most beastly stuff. It is that sticky sweet kind.</p>

<p>JOHN \direct{who is perhaps still playing at parents}¤
Never mind, father, it will soon be over.</p>

<blockquote><i>(A spasm of ill-will to</i> JOHN <i>cuts through</i> MR. DARLING, <i>and is gone.</i> WENDY <i>returns panting.)</i></blockquote>

<p>WENDY. Here it is, father; I have been as quick as I could.</p>

<p>MR. DARLING \direct{with a sarcasm that is completely thrown away on her}¤
You have been wonderfully quick, precious quick!</p>

<blockquote><i>(He is now at the foot of</i> MICHAEL'S <i>bed,</i> NANA <i>is by its side, holding the medicine spoon insinuatingly in her mouth.)</i></blockquote>

<p>WENDY \direct{proudly, as she pours out</i> MR. DARLING'S <i>medicine}¤
Michael, now you will see how father takes it.</p>

<p>MR. DARLING \direct{hedging}¤
Michael first.</p>

<p>MICHAEL \direct{full of unworthy suspicions}¤
Father first.</p>

<p>MR. DARLING. It will make me sick, you know.</p>

<p>JOHN \direct{lightly}¤
Come on, father.</p>

<p>MR. DARLING. Hold your tongue, sir.</p>

<p>WENDY \direct{disturbed}¤
I thought you took it quite easily, father, saying 'Thank you, kind parents, for&mdash;&mdash;&mdash;'</p>

<p>MR. DARLING. That is not the point; the point is that there is more in my glass than in Michael's spoon. It isn't fair, I swear though it were with my last breath, it is not fair.</p>

<p>MICHAEL \direct{coldly}¤
Father, I'm waiting.</p>

<p>MR. DARLING. It's all very well to say you are waiting; soam I waiting.</p>

<p>MICHAEL. Father 's a cowardy custard.</p>

<p>MR. DARLING. So are you a cowardy custard.</p>

<blockquote><i>(They are now glaring at each other.)</i></blockquote>

<p>MICHAEL. I am not frightened.</p>

<p>MR. DARLING. Neither am I frightened.</p>

<p>MICHAEL. Well, then, take it.</p>

<p>MR. DARLING. Well, then, you take it.</p>

<p>WENDY \direct{butting in again}¤
Why not take it at the same time?</p>

<p>MR. DARLING \direct{haughtily}¤
Certainly. Are you ready, Michael?</p>

<p>WENDY \direct{as nothing has happened}¤
One&mdash;two&mdash;three.</p>

<blockquote>(MICHAEL <i>partakes, but</i> MR. DARLING <i>resorts to hanky-panky.)</i></blockquote>

<p>JOHN. Father hasn't taken his!</p>

<blockquote>(MICHAEL <i>howls.)</i></blockquote>

<p>WENDY \direct{inexpressibly pained}¤
Oh father!</p>

<p>MR. DARLING <i>(who has been hiding the glass behind him).</i>What do you mean by 'oh father'? Stop that row, Michael. I meant to take mine but I&mdash;missed it. (NANA <i>shakes her head sadly over him, and goes into the bathroom. They are all looking as if they did not admire him, and nothing so dashes a temperamental man.)</i> I say, I have just thought of a splendid joke.
\direct{They brighten.}
I shall pour my medicine into Nana's bowl, and she will drink it thinking it is milk! <i>The pleasantry does not appeal, but he prepares the joke, listening for appreciation.)</i></p>

<p>WENDY. Poor darling Nana!</p>

<p>MR. DARLING. You silly little things; to your beds everyone of you; I am ashamed of you.</p>

<blockquote><i>(They steal to their beds as</i> MRS. DARLING <i>returns with the chocolate.)</i></blockquote>

<p>MRS. DARLING. Well, is it all over?</p>

<p>MICHAEL. Father didn't&mdash;&mdash;<i>(Father glares.)</i></p>

<p>MR. DARLING. All over, dear, quite satisfactorily. (NANA <i>comes back.)</i> Nana, good dog, good girl; I have put a little milk into your bowl.
\direct{The bowl is by the kennel, and</i> NANA <i>begins to lap, only begins. She retreats into the kennel.}</p>

<p>MRS. DARLING. What is the matter, Nana?</p>

<p>MR. DARLING \direct{uneasily}¤
Nothing, nothing.</p>

<p>MRS. DARLING \direct{smelling the bowl}¤
George, it is your medicine!</p>

<blockquote><i>(The children break into lamentation. He gives his wife an imploring look; he is begging for one smile, but does not get it. In consequence he goes from bad to worse.)</i></blockquote>

<p>MR. DARLING. It was only a joke. Much good my wearing myself to the bone trying to be funny in this house.</p>

<p>WENDY \direct{on her knees by the kennel}¤
Father, Nana is crying.</p>

<p>MR. DARLING. Coddle her; nobody coddles me. Oh dear no. I am only the bread-winner, why should I be coddled? Why, why, why?</p>

<p>MRS. DARLING. George, not so loud; the servants will hear you.</p>

<blockquote><i>(There is only one maid, absurdly small too, but they have got into the way of calling her the servants.)</i></blockquote>

<p>MR. DARLING \direct{defiant}¤
Let them hear me; bring in the whole world. ( <i>The desperate man, who has not been in fresh air for days, has now lost all self-control.)</i> I refuse to allow that dog to lord it in my nursery for one hour longer. (NANA <i>supplicates him.)</i> In vain, in vain, the proper place for you is the yard, and there you go to be tied up this instant.</p>

<blockquote>(NANA <i>again retreats into the kennel, and the children add their prayers to hers.)</i></blockquote>

<p>MRS. DARLING \direct{who knows how contrite he will be for this presently}¤
George, George, remember what I told you about that boy.</p>

<p>MR. DARLING. Am I master in this house or is she?
\direct{To</i> NANA <i>fiercely}¤
Come along.
\direct{He thunders at her, but she indicates that she has reasons not worth troubling him with for remaining where she is. He resorts to a false bonhomie.}
There, there, did she think he was angry with her, poor Nana?
\direct{She wriggles a response in the affirmative.}
Good Nana, pretty Nana.
\direct{She has seldom been called pretty, and it has the old effect. She plays rub-a-dub with her paws, which is how a dog blushes.}
She will come to her kind master, won't she? won't she?
\direct{She advances, retreats, waggles her head, her tail, and eventually goes to him. He seizes her collar in an iron grip and amid the cries of his progeny drags her from the room. They listen, for her remonstrances are not inaudible.}</p>

<p>MRS. DARLING. Be brave, my dears.</p>

<p>WENDY. He is chaining Nana up!</p>

<blockquote><i>(This unfortunately is what he is doing, though we cannot see him. Let us hope that he then retires to his study, looks up the word 'temper' in his</i> Thesaurus, <i>and under the influence of those benign pages becomes a better man. In the meantime the children have been put to bed in unwonted silence, and</i> MRS. DARLING <i>lights the night-lights over the beds.)</i></blockquote>

<p>JOHN \direct{as the barking below goes on}¤
She is awfully unhappy.</p>

<p>WENDY. That is not Nana's unhappy bark. That is her bark when she smells danger.</p>

<p>MRS. DARLING \direct{remembering that boy}¤
Danger! Are you sure, Wendy?</p>

<p>WENDY \direct{the one of the family, for there is one in every family, who can be trusted to know or not to know}¤
Oh yes.</p>

<blockquote><i>(Her mother looks this way and that from the window.)</i></blockquote>

<p>JOHN. Is anything there?</p>

<p>MRS. DARLING. All quite quiet and still. Oh, how I wish I was not going out to dinner to-night.</p>

<p>MICHAEL. Can anything harm us, mother, after the night-lights are lit?</p>

<p>MRS. DARLING. Nothing precious. They are the eyes amother leaves behind her to guard her children.</p>

<blockquote><i>(Nevertheless we may be sure she means to tell</i> LIZA, <i>the little maid, to look in on them frequently till she comes home. She goes from bed to bed, after her custom, tucking them in and crooning a lullaby.)</i></blockquote>

<p>MICHAEL \direct{drowsily}¤
Mother, I 'm glad of you.</p>

<p>MRS. DARLING \direct{with a last look round, her hand on the switch}¤
Dear night-lights that protect my sleeping babes, burn clear and steadfast to-night.</p>

<blockquote><i>(The nursery darkens and she is gone, intentionally leaving the door ajar. Something uncanny is going to happen, we expect, for a quiver has passed through the room, just sufficient to touch the night-lights. They blink three times one after the other and go out, precisely as children (whom familiarity has made them resemble) fall asleep. There is another light in the room now, no larger than</i> MRS. DARLING'S <i>fist, and in the time we have taken to say this it has been into the drawers and wardrobe and searched pockets, as it darts about looking for a certain shadow. Then the window is blown open, probably by the smallest and therefore most mischievous star, and</i> PETER PAN <i>flies into the room. In so far as he is dressed at all it is in autumn leaves and cobwebs.)</i></blockquote>

<p>PETER (in <i>a whisper).</i> Tinker Bell, Tink, are you there?
\direct{A jug lights up.}
Oh, do come out of that jug. (TINK<i>flashes hither and thither?)</i> Do you know where they put it?
\direct{The answer comes as of a tinkle of bells; it is the fairy language.</i> PETER <i>can speak it, but it bores him.}
Which big box? This one? But which drawer? Yes, do show me. (TINK <i>pops into the drawer where the shadow is, but before</i>PETER <i>can reach it,</i> WENDY <i>moves in her sleep. He flies onto the mantelshelf as a hiding-place. Then, as she has not waked, he flutters over the beds as an easy way to observe the occupants, closes the window softly, wafts himself to the drawer and scatters its contents to the floor, as kings on their wedding day toss ha'pence to the crowd. In his joy at finding his shadow he forgets that he has shut up</i> TINK <i>in the drawer. He sits on the floor with the shadow, confident that he and it will join like drops of water. Then he tries to stick it on with soap from the bathroom, and this failing also, he subsides dejectedly on the floor. This wakens</i> WENDY, <i>who sits up, and is pleasantly interested to see a stranger.)</i></p>

<p><i>WENDY (courteously).</i> Boy, why are you crying?</p>

<blockquote><i>(He jump up, and crossing to the foot of the bed bows to her in the fairy way.</i> WENDY, <i>impressed, bows to him from the bed.)</i></blockquote>

<p>PETER. What is your name?</p>

<p>WENDY \direct{well satisfied}¤
Wendy Moira Angela Darling.What is yours?</p>

<p>PETER \direct{finding it lamentably brief}¤
Peter Pan.</p>

<p>WENDY. Is that all?</p>

<p>PETER \direct{biting his lip}¤
Yes.</p>

<p>WENDY \direct{politely}¤
I am so sorry.</p>

<p>PETER. It doesn't matter.</p>

<p>WENDY. Where do you live?</p>

<p>PETER. Second to the right and then straight on till morning.</p>

<p>WENDY. What a funny address!</p>

<p>PETER. No, it isn't.</p>

<p>WENDY. I mean, is that what they put on the letters?</p>

<p>PETER. Don't get any letters.</p>

<p>WENDY. But your mother gets letters?</p>

<p>PETER. Don't have a mother.</p>

<p>WENDY. Peter!</p>

<blockquote><i>(She leaps out of bed to put her arms round him, but he draws back; he does not know why, but he knows he must draw back.)</i></blockquote>

<p>PETER. You mustn't touch me.</p>

<p>WENDY. Why?</p>

<p>PETER. No one must ever touch me.</p>

<p>WENDY. Why?</p>

<p>PETER. I don't know.</p>

<blockquote><i>(He is never touched by any one in the play.)</i></blockquote>

<p>WENDY. No wonder you were crying.</p>

<p>PETER. I wasn't crying. But I can't get my shadow to stick on.</p>

<p>WENDY. It has come off! How awful.
\direct{Looking at the spot where he had lain.}
Peter, you have been trying to stick it on with soap!</p>

<p>PETER \direct{snappily}¤
Well then?</p>

<p>WENDY. It must be sewn on.</p>

<p>PETER. What is 'sewn'?</p>

<p>WENDY. You are dreadfully ignorant.</p>

<p>PETER. No, I 'm not.</p>

<p>WENDY. I will sew it on for you, my little man. But we must have more light.
\direct{She touches something, and to his astonishment the room is illuminated.}
Sit here. I dare say it will hurt a little.</p>

<p>PETER \direct{a recent remark of hers rankling}¤
I never cry.
\direct{She seems to attach the shadow. He tests the combination.}
It isn't quite itself yet.</p>

<p>WENDY. Perhaps I should have ironed it.
\direct{It awakes and is as glad to be back with him as he to have it. He and his shadow dance together. He is showing off now. He crows</i> <i>like a cock. He would fly in order to impress</i> WENDY <i>further</i> <i>if he knew that there is anything unusual in that.}</p>

<p>PETER. Wendy, look, look; oh the cleverness of me!</p>

<p>WENDY. You conceit, of course I did nothing!</p>

<p>PETER. You did a little.</p>

<p>WENDY \direct{wounded}¤
A little! If I am no use I can at least withdraw.</p>

<blockquote><i>(With one haughty leap she is again in bed with the sheet over her face. Popping on to the end of the bed the artful one appeals.)</i></blockquote>

<p>PETER. Wendy,. don't withdraw. I can't help crowing, Wendy, when I'm pleased with myself. Wendy, one girl is worth more than twenty boys.</p>

<p>WENDY \direct{peeping over the sheet}¤
You really think so, Peter?</p>

<p>PETER. Yes, I do.</p>

<p>WENDY. I think it's perfectly sweet of you, and I shall get up again.
\direct{They sit together on the side of the bed.}
I shall give you a kiss if you like.</p>

<p>PETER. Thank you.
\direct{He holds out his hand.}</p>

<p>WENDY \direct{aghast}¤
Don't you know what a kiss is?</p>

<p>PETER. I shall know when you give it me.
\direct{Not to hurt his feelings she gives him her thimble.}
Now shall I give youa kiss?</p>

<p>WENDY \direct{primly}¤
If you please.
\direct{He pulls an acorn button off his person and bestows it on her. She is shocked but considerate.}
I will wear it on this chain round my neck. Peter, how old are you?</p>

<p>PETER \direct{blithely}¤
I don't know, but quite young, Wendy. I ran away the day I was born.</p>

<p>WENDY. Ran away, why?</p>

<p>PETER. Because I heard father and mother talking of what I was to be when I became a man. I want always to be a little boy and to have fun; so I ran away to Kensington Gardens and lived a long time among the fairies.</p>

<p>WENDY \direct{with great eyes}¤
You know fairies, Peter!</p>

<p>PETER \direct{surprised that this should be a recommendation}¤
Yes, but they are nearly all dead now.
\direct{Baldly}¤
You see, Wendy, when the first baby laughed for the first time, the laugh broke into a thousand pieces and they all went skipping about, and that was the beginning of fairies. And now when every new baby is born its first laugh becomes a fairy. So there ought to be one fairy for every boy or girl,</p>

<p>WENDY \direct{breathlessly}¤
Ought to be? Isn't there?</p>

<p>PETER. Oh no. Children know such a lot now. Soon they don't believe in fairies, and every time a child says 'I don't believe in fairies' there is a fairy somewhere that falls down dead.
\direct{He skips about heartlessly.}</p>

<p>WENDY. Poor things!</p>

<p>PETER. (<i>to whom this statement recalls a forgotten friend)</i>. I can't think where she has gone. Tinker Bell, Tink, where are you?</p>

<p>WENDY \direct{thrilling}¤
Peter, you don't mean to tell me that there is a fairy in this room!</p>

<p>PETER (<i>flitting about in search</i>). She came with me. You don't hear anything, do you?</p>

<p>WENDY. I hear&mdash;the only sound I hear is like a tinkle of bells.</p>

<p>PETER. That is the fairy language. I hear it too.</p>

<p>WENDY. It seems to come from over there.</p>

<p>PETER. (<i>with shameless glee.</i>) Wendy, I believe I shut her up in that drawer!</p>

<blockquote><i>(He releases</i> TINK, <i>who darts about in a fury using language it is perhaps as well we don't understand.)</i></blockquote>

You needn't say that; I 'm very sorry, but how could I know you were in the drawer? 

<p>WENDY \direct{her eyes dancing in pursuit of the delicious creature}¤
Oh, Peter, if only she would stand still and let me see her!</p>

<p>PETER \direct{indifferently}¤
They hardly ever stand still.</p>

<blockquote><i>(To show that she can do even this</i> TINK <i>pauses between two ticks of the cuckoo clock.)</i></blockquote>

<p>WENDY. I see her, the lovely! where is she now?</p>

<p>PETER. She is behind the clock. Tink, this lady wishes you were her fairy.
\direct{The answer comes immediately.}</p>

<p>WENDY. What does she say?</p>

<p>PETER. She is not very polite. She says you are a great ugly girl, and that she is my fairy. You know, Tink, you can't be my fairy because I am a gentleman and you are a lady.</p>

<blockquote> <p>(TINK <i>replies.)</i></p> </blockquote>

<p>WENDY. What did she say?</p>

<p>PETER. She said 'You silly ass.' She is quite a common girl, you know. She is called Tinker Bell because she mends the fairy pots and kettles.</p>

<blockquote>(<i>They have reached a chair,</i> WENDY <i>in the ordinary way and</i> PETER <i>through a hole in the back.)</i></blockquote>

<p>WENDY. Where do you live now?</p>

<p>PETER. With the lost boys.</p>

<p>WENDY. Who are they?</p>

<p>PETER. They are the children who fall out of their prams when the nurse is looking the other way. If they are not claimed in seven days they are sent far away to the Never-Land. I 'm captain.</p>

<p>WENDY. What fun it must be.</p>

<p>PETER \direct{craftily}¤
Yes, but we are rather lonely. You see, Wendy, we have no female companionship.</p>

<p>WENDY. Are none of the other children girls?</p>

<p>PETER. Oh no; girls, you know, are much too clever to fall out of their prams.</p>

<p>WENDY. Peter, it is perfectly lovely the way you talk about girls. John there just depises us.<br> (PETER, <i>for the first time, has a good look at</i> JOHN. <i>He then neatly tumbles him out of bed.)</i></p>

<p>You wicked! you are not captain here.
\direct{She bends over her brother who is prone on the floor.}
After all he hasn't wakened, and you meant to be kind.
\direct{Having now done her duty she forgets</i> JOHN, <i>who blissfully sleep on.}
Peter, you may give me a kiss.</p>

<p>PETER \direct{cynically}¤
I thought you would want it back.
\direct{He offers her the thimble.}</p>

<p>WENDY \direct{artfully}¤
Oh dear, I didn't mean a kiss, Peter. I meant a thimble.</p>

<p>PETER <i>(only half placated).</i>What is that?</p>

<p>WENDY. It is like this.
\direct{She leans forward to give a demonstration, but something prevents the meeting of their faces.}</p>

<p>PETER \direct{satisfied}¤
Now shall I give you a thimble?</p>

<p>WENDY. If you please.
\direct{Before he can even draw near she screams.}</p>

<p>PETER. What is it?</p>

<p>WENDY. It was exactly as if some one were pulling my hairl</p>

<p>PETER. That must have been Tink. I never knew her so naughty before.</p>

<blockquote>(TINK <i>speaks. She is in the jug again.)</i></blockquote>

<p>WENDY. What does she say?</p>

<p>PETER. She says she will do that every time I give you a thimble.</p>

<p>WENDY. But why?</p>

<p>PETER \direct{equally nonplussed}¤
Why, Tink?
\direct{He has to translate the answer.}
She said 'You silly ass' again.</p>

<p>WENDY. She is very impertinent.
\direct{They are sitting on the floor now.}
Peter, why did you come to our nursery window?</p>

<p>PETER. To try to hear stories None of us knows any stories.</p>

<p>WENDY. How perfectly awful!</p>

<p>PETER. Do you know why swallows build in the eaves of houses? It is to listen to the stories. Wendy, your mother was telling you such a lovely story.</p>

<p>WENDY. Which story was it?</p>

<p>PETER. About the prince, and he couldn't find the lady who wore the glass slipper.</p>

<p>WENDY. That was Cinderella. Peter, he found her and they were happy ever after.</p>

<p>PETER. I am glad.
\direct{They have worked their way along the floor close to each other, but he now jumps up.}</p>

<p>WENDY. Where are you going?</p>

<p>PETER \direct{already on his way to the window}¤
To tell the other boys.</p>

<p>WENDY. Don't go, Peter. I know lots of stories. The stories I could tell to the boys!</p>

<p>PETER \direct{gleaming}¤
Come on! We'll fly.</p>

<p>WENDY. Fly? You can fly!</p>

<blockquote><i>(How he would like to rip those stories out of her; he is dangerous now.)</i></blockquote>

<p>PETER. Wendy, come with me.</p>

<p>WENDY. Oh dear, I mustn't. Think of mother. Besides, I can't fly.</p>

<p>PETER. I'll teach you.</p>

<p>WENDY. How lovely to fly!</p>

<p>PETER. I'll teach you how to jump on the wind's back and then away we go. Wendy, when you are sleeping in your silly bed you might be flying about with me, saying funny things tothe stars. There are mermaids, Wendy, with long tails.
\direct{She just succeeds in remaining on the nursery floor.}
Wendy, how we should all respect you.</p>

<blockquote><i>(At this she strikes her colours.)</i></blockquote>

<p>WENDY. Of course it's awfully fascinating! Would you teach John and Michael to fly too?</p>

<p>PETER \direct{indifferently}¤
If you like.</p>

<p>WENDY <i>(playing rum-turn on</i> JOHN). John, wake up; there is a boy here who is to teach us to fly.</p>

<p>JOHN. Is there? Then I shall get up.
\direct{He raises his headfrom the floor.}
Hullo, I am up!</p>

<p>WENDY. Michael, open your eyes. This boy is to teach us to fly.</p>

<blockquote><i>(The sleepers are at once as awake as their father's razor;but before a question can be asked</i> NANA'S <i>bark is heard.)</i></blockquote>

<p>JOHN. Out with the light, quick, hide!</p>

<blockquote><i>(When the maid</i> LIZA, <i>who is so small that when she says she will never see ten again one can scarcely believe her, enters with a firm hand on the troubled</i> NANA'S <i>chain the room is in comparative darkness.)</i></blockquote>

<p>LIZA. There, you suspicious brute, they are perfectly safe, aren't they? Every one of the little angels sound asleep in bed. Listen to their gentle breathing. (NANA'S <i>sense of smell here helps to her undoing instead of hindering it. She knows that they are in the room.</i> MICHAEL, <i>who is behind the window curtain, is so encouraged by</i> LIZA'S <i>last remark that he breathes too loudly.</i> NANA <i>knows that kind of breathing and tries to break from her keeper's control.)</i> No more of it, Nana.
\direct{Wagging a finger at her}¤
I warn you if you bark again I shall go straight for master and missus and bring them home from the party, and then won't master whip you just! Come along, you naughty dog.</p>

<blockquote> <p><i>(The unhappy</i> NANA is <i>led away. The children emerge exulting from their various hiding-places. In their brief absence from the scene strange things have been done to them; but it is not for us to reveal a mysterious secret of the stage. They look just the same.)</i></p> </blockquote>

<p>JOHN. I say, can you really fly.</p>

<p>PETER. Look!
\direct{He is now over their heads.}
</p>

<p>WENDY. Oh, how sweet!</p>

<p>PETER. I 'm sweet, oh, I am sweet!</p>

<blockquote><i>(It looks so easy that they try it first from the floor <b>and</b>then from their beds, without encouraging results.)</i></blockquote>

<p>JOHN \direct{rubbing his knees}¤
How do you do it?</p>

<p>PETER \direct{descending}¤
You just think lovely wonderful thoughts and they lift you up in the air.
\direct{He is off again.}
</p>

<p>JOHN. You are so nippy at it; couldn't you do it very slowly once? (PETER <i>does it slowly.)</i> I 've got it now, Wendy.
\direct{He tries; no, he has not got it, poor stay-at-home, though he knows the names of all the counties in England and</i> PETER <i>does not know one.}
</p>

<p>PETER. I must blow the fairy dust on you first.
\direct{Fortunately his garments are smeared with it and he blows some dust on each.}
Now, try; try from the bed. Just wiggle your shoulders this way, and then let go.</p>

<blockquote><i>(The gallant</i> MICHAEL <i>is the first to let go, and is borne across the room.)</i></blockquote>

<p>MICHAEL <i>(with a yell that should have disturbed</i> LIZA). I flewed!</p>

<blockquote>(JOHN <i>lets go, and meets</i> WENDY <i>near the bathroom door though they had both aimed in an opposite direction.)</i></blockquote>

<p>WENDY. Oh, lovely!</p>

<p>JOHN \direct{tending to be upside down}¤
How ripping!</p>

<p>MICHAEL \direct{playing whack on a chair}¤
I do like it!</p>

<p>THE THREE. Look at me, look at me, look at me!</p>

<blockquote><i>(They are not nearly so elegant in the air as</i> PETER, <i>but their heads have bumped the ceiling, and there is nothing more delicious than that.)</i></blockquote>

<p>JOHN \direct{who can even go backwards}¤
I say, why shouldn't we go out?</p>

<p>PETER. There are pirates.</p>

<p>JOHN. Pirates!
\direct{He grabs his tall Sunday hat.}
Let us go at once!</p>

<blockquote>(TINK <i>does not like it. She darts at their hair. From down below in the street the lighted window must present an unwonted spectacle: the shadows of children revolving in the room like a merry-go-round. This is perhaps what</i> MR. <i>and</i> MRS. DARLING <i>see as they come hurrying home from the party, brought by</i> NANA <i>who, you may be sure, has broken her chain.</i> PETER'S <i>accomplice, the little star, has seen them coming, and again the window blows open.)</i></blockquote>

<p>PETER \direct{as if he had heard the star whisper 'Cave'}¤
Now come!</p>

<blockquote> <p><i>(Breaking the circle he flies out of the window over the trees of the square and over the house-tops, and the others follow like a flight of birds. The broken-hearted father and mother arrive just in time to get a nip from TINK as she too sets out for the Never Land.)</i></p> </blockquote>
