% !TEX program = pdflatex
% !TEX encoding = UTF-8
% !TEX spellcheck = en_GB
% !TEX root = peter-pan.tex

\Act{The Never Land}

\endinput

<p><i>When the blind goes up all is so dark that you scarcely know it has gone up. This is because if you were to see the island bang</i> (<i>as Peter would say) the wonders of it might hurt your eyes. If you all came in spectacles perhaps you could see it bang, but to make a rule of that kind would be a pity. The first thing seen is merely some whitish dots trudging along the sward, and you can guess from their tinkling that they are probably fairies of the commoner sort going home afoot from some party and having a cheery tiff by the way. Then Peter's star wakes up, and in the blink of it, which is much stronger than in our stars, you can make out masses of trees, and you think you see wild beasts stealing past to drink, though what you see is not the beasts themselves but only the shadows of them. They are really out pictorially to greet Peter in the way they think he would like them to greet him; and for the same reason the mermaids basking in the lagoon beyond the trees are carefully combing their hair; and for the same reason the pirates are landing invisibly from the longboat, invisibly to you but not to the redskins, whom none can see or hear because they are on the war-path. The whole island, in short, which has been having a slack time in Peter's absence, is now in a ferment because the tidings has leaked out that he is on his way back; and everybody and everything know that they will catch it from him if they don't give satisfaction. While you have been told this the sun (another of his servants) has been bestirring himself. Those of you who may have thought it wiser after all to begin this Act in spectacles may now take them off.</i></p>

<p><i>What you see is the Never Land. You have often half seen it before, or even three-quarters, after the night-lights were lit, and you might then have beached your coracle on it if you had not always at the great moment fallen asleep. I dare say you have chucked things on to it, the things you can't find in the morning. In the daytime you think the Never Land is only make-believe, and so it is to the likes of you, but this is the Never Land come true. It is an open-air scene, a forest, with a beautiful lagoon beyond but not really far away, for the Never Land is very compact, not large and sprawly with tedious distances between one adventure and another, but nicely crammed. It is summer time on the trees and on the lagoon but winter on the river, which is not remarkable on Peter's island where all the four seasons may pass while you are filling a jug at the well. Peter's home is at this very spot, but you could not point out the way into it even if you were told which is the entrance, not even if you were told that there are seven of them. You know now because you have just seen one of the lost boys emerge. Theholes in these seven great hollow trees are the 'doors' down to Peter's home, and he made seven because, despite his cleverness, he thought seven boys must need seven doors.</i></p>

<p><i>The boy who has emerged from his tree is Slightly, who has perhaps been driven from the abode below by companions less musical than himself. Quite possibly a genius Slightly has with him his home-made whistle to which he capers entrancingly, with no audience save a Never ostrich which is also musically inclined. Unable to imitate Slightly's graces the bird falls so low as to burlesque them and is driven from the entertainment. Other lost boys climb up the trunks or drop from branches, and now we see the six of them, all in the skins of animals they think they have shot, and so round and furry in them that if they fall they roll. Tootles is not the least brave though the most unfortunate of this gallant band. He has been in fewer adventures than any of them because the big things constantly happen while he has stepped round the corner; he will go off, for instance, in some quiet hour to gather firewood, and then when he returns the others will be sweeping up the blood. Instead of souring his nature this has sweetened it and he is the humblest of the band. (Nibs is more gay and debonair, Slightly more conceited. Slightly thinks he remembers the days before he was lost, with their manners and customs.</i> <i>Curly is a pickle, and so often has he had to deliver up his person when Peter said sternly, 'Stand forth the one who did this thing,' that now he stands forth whether he has done it or not. The other two are First Twin and Second Twin, who cannot be described because we should probably be describing the wrong one. Hunkering on the ground or peeking out of their holes, the six are not unlike village gossips gathered round the pump.</i></p>

<p>TOOTLES. Has Peter come back yet, Slightly?</p>

<p>SLIGHTLY <i>(with a solemnity that he thinks suits the occasion).</i> No, Tootles, no.</p>

<blockquote><i>(They are like dogs waiting for the master to tell them that the day has begun.)</i></blockquote>

<p>CURLY <i>(as if Peter might be listening).</i> I do wish he would come back.</p>

<p>TOOTLES. I am always afraid of the pirates when Peter is not here to protect us.</p>

<p>SLIGHTLY. I am not afraid of pirates. Nothing frightens me. But I do wish Peter would come back and tell us whether he has heard anything more about Cinderella.</p>

<p>SECOND TWIN <i>(with diffidence).</i> Slightly, I dreamt last night that the prince found Cinderella.</p>

<p>FIRST TWIN <i>(who is intellectually the superior of the two).</i> Twin, I think you should not have dreamt that, for I didn't, and Peter may say we oughtn't to dream differently, being twins, you know.</p>

<p>TOOTLES. I am awfully anxious about Cinderella. You see, not knowing anything about my own mother I am fond of thinking that she was rather like Cinderella.</p>

<blockquote><i>(This is received with derision.)</i></blockquote>

<p>NIBS. All I remember about my mother is that she often said to father, 'Oh how I wish I had a cheque book of my own.' I don't know what a cheque book is, but I should just love to give my mother one.</p>

<p>SLIGHTLY <i>(as usual).</i> My mother was fonder of me than your mothers were of you. <i>(Uproar.)</i> Oh yes, she was. Peter had to make up names for you, but my mother had wrote my name on the pinafore I was lost in. 'Slightly Soiled'; that's my name.</p>

<blockquote><i>(They fall upon him pugnaciously</i>; <i>not that they are really worrying about their mothers, who are now as important to them as a piece of string, but because any excuse is good enough for a shindy. Not for long is he belaboured, for a sound is heard that sends them scurrying down their holes; in a second of time the scene is bereft of human life. What they have heard from near-by is a verse of the dreadful song with which on the Never Land the prates stealthily trumpet their approach&mdash;</i></blockquote>

<blockquote> <p>Yo ho, yo ho, the pirate life,</p>

<p>The flag of skull and bones,</p>

<p>A merry hour, a hempen rope,</p>

<p>And hey for Davy Jones!</p> </blockquote>

<blockquote> <p><i>The pirates appear upon the frozen river dragging a raft, on which reclines among cushions that dark and fearful man,</i> CAPTAIN JAS. HOOK. <i>A more villainous-looking brotherhood of men never hung in a row on Execution dock. Here, his great arms bare, pieces of eight in his ears as ornaments, is the handsome</i> CECCO, <i>who cut his name on the back of the governor of the prison at Gao. Heavier in the pull is the gigantic black who has had many names since the first one terrified dusky children on the banks of the</i> Guidjo-mo. BILL JUKES com<i>es next, every inch of him tattooed, the same</i> JUKES <i>who got six dozen on the</i> Walrus <i>from</i> FLINT. <i>Following these</i> are COOKSON, <i>said to be</i> BLACK. MURPHY'S <i>brother (but this was never proved);</i> and GENTLEMAN STARKEY, <i>once an usher in a school; and</i> SKYLIGHTS <i>(Morgan's Skylights); and</i> NOODLER, <i>whose hands are fixed on backwards; and the spectacled boatswain,</i> SMEE, <i>the only Nonconformist in</i> HOOK'S <i>crew; and other ruffians long known and feared on the Spanish main.</i></p>

<p><i>Cruelest jewel in that dark setting is</i> HOOK <i>himself, cadaverous and blackavised, his hair dressed in long curls which look like black candles about to melt, his eyes blue as the forget-me-not and of a profound insensibility, save when he claws, at which time a red spot appears in them. He has an iron hook instead of a right hand, and it is with this he claws. He is never more sinister than</i> <i>when he is most polite, and the elegance of his diction, the distinction of his demeanour, show him one of a different class from his crew, a solitary among uncultured companions. This courtliness impresses even his victims on the high seas, who note that he always says 'Sorry' when prodding them along the flank. A man of indomitable courage, the only thing at which he flinches is the sight of his own blood, which is thick and of an unusual colour. At his public school they said of him that he 'bled yellow.' In dress he apes the dandiacal associated with Charles II., having heard it said in an earlier period of his career that he bore a strange resemblance to the ill-fated Stuarts. A holder of his own contrivance is in his mouth enabling him to smoke two cigars at once. Those, however, who have seen him in the flesh, which is an inadequate term for his earthly tenement, agree that the grimmest part of him is his iron claw.</i></p>

<p><i>They continue their distasteful singing as they disembark&mdash;</i></p> </blockquote>

<blockquote> <p>Avast, belay, yo ho, heave to,</p>

<p>A-pirating we go,</p>

<p>And if we 're parted by a shot</p>

<p>We 're sure to meet below!</p> </blockquote>

<blockquote> <p>NIBS, <i>the only one of the boys who has not sought safety in his tree, is seen for a moment near the lagoon, and</i> STARKEY'S <i>pistol is at once upraised. The captain twists his hook in him.)</i></p> </blockquote>

<p>STARKEY <i>(abject).</i> Captain, let go</p>

<p>!HOOK. Put back that pistol, first.</p>

<p>STARKEY. 'Twas one of those boys you hate; I could haveshot him dead.</p>

<p>HOOK. Ay, and the sound would have brought Tiger Lily's redskins on us. Do you want to lose your scalp?</p>

<p>SMEE <i>(wriggling his cutlass pleasantly).</i> That is true. Shall I after him, Captain, and tickle him with Johnny Corkscrew? Johnny is a silent fellow.</p>

<p>HOOK. Not now. He is only one, and I want to mischief all the seven. Scatter and look for them. <i>(The boatswain whistles his instructions, and the men disperse on their frightful errand. With none to hear save</i> SMEE, HOOK <i>becomes confidential.)</i> Most of all I want their captain, Peter Pan. 'Twas he cut off my arm. I have waited long to shake his hand with this. <i>(Luxuriating.)</i> Oh, I 'll tear him!</p>

<p>SMEE <i>(always ready for a chat).</i> Yet I have oft heard you say your hook was worth a score of hands, for combing the hair and other homely uses.</p>

<p>HOOK. If I was a mother I would pray to have my children born with this instead of that <i>(his left arm creeps nervously behind him. He has a galling remembrance).</i> Smee, Pan flung my arm to a crocodile that happened to be passingby.</p>

<p>SMEE. I have often noticed your strange dread of crocodiles.</p>

<p>HOOK <i>(pettishly).</i> Not of crocodiles but of that one crocodile. <i>(He lays bare a lacerated heart.)</i> The brute liked my arm so much, Smee, that he has followed me ever since, from sea to sea, and from land to land, licking his lips for the rest of me.</p>

<p>SMEE <i>(looking for the bright side).</i> In a way it is a sort of compliment.</p>

<p>HOOK <i>(with dignity).</i> I want no such compliments; I want Peter Pan, who first gave the brute his taste for me. Smee, that crocodile would have had me before now, but by a lucky chance he swallowed a clock, and it goes tick, tick, tick, tick inside him; and so before he can reach me I hear the tick and bolt. <i>(He emits a hollow rumble.)</i> Once I heard it strike six within him.</p>

<p>SMEE <i>(sombrely).</i> Some day the clock will run down,and then he'll get you.</p>

<p>HOOK <i>(a broken man).</i> Ay, that is the fear that haunts me.<i>(He rises.)</i> Smee, this seat is hot; odds, bobs, hammer and tongs, I am burning.</p>

<blockquote> <p><i>(He has been sitting, he thinks, on one of the island mushrooms, which are of enormous size. But this is a hand-painted one placed here in times of danger to conceal a chimney. They remove it, and tell-tale smoke issues; also, alas, the sound of children's voices.)</i></p> </blockquote>

<p>SMEE. A chimney!</p>

<p>HOOK <i>(avidly).</i> Listen! Smee, 'tis plain they live here, beneath the ground. <i>(He replaces the mushroom. His brain works tortuously.)</i></p>

<p>SMEE <i>(hopefully).</i> Unrip your plan, Captain.</p>

<p>HOOK. To return to the boat and cook a large rich cakeof jolly thickness with sugar on it, green sugar. There can be but one room below, for there is but one chimney. The silly moles had not the sense to see that they did not need a door apiece. We must leave the cake on the shore of the mermaids' lagoon. These boys are always swimming about there, trying to catch the mermaids. They will find the cake and gobble it up, because, having no mother, they don't know how dangerous 'tis to eat rich damp cake. They will die!</p>

<p>SMEE <i>(fascinated).</i> It is the wickedest, prettiest policy ever I heard of,</p>

<p>HOOK <i>(meaning well).</i> Shake hands on 't.</p>

<p>SMEE. No, Captain, no.</p>

<blockquote><i>(He has to link with the hook, but he does not join in the song.)</i></blockquote>

<p>HOOK. Yo ho, yo ho, when I say 'paw,'</p>

<blockquote> By fear they're overtook,<br> Naught's left upon your bones when you.<br> Have shaken hands with Hook!<br> <br> <i>(Frightened by a tug at his hand,</i> SMEE <i>is joining in the chorus when another sound stills them both. It is a tick, tick as of a clock, whose significance</i> HOOK <i>is, naturally, the first to recognise, 'The crocodile!' he cries, and totters from the scene.</i> SMEE <i>follows. A huge crocodile, of one thought compact, passes across, ticking, and oozes after them. The wood is now so silent that you may be sure it is full of redskins.</i> TIGER LILY <i>comes first. She is the belle of the Piccaninny tribe, whose braves would all have her to wife, but she wards them off witha hatchet. She puts her ear to the ground and listens, then beckons, and</i> GREAT BIG LITTLE PANTHER <i>and the tribe are around her, carpeting the ground. Far away some one treads on a dry leaf.')</i></blockquote>

<p>TIGER LILY. Pirates! <i>(They do not draw their knives) the knives slip into their hands.)</i> Have um scalps? What you say?</p>

<p>PANTHER. Scalp um, oho, velly quick.</p>

<p>THE BRAVES <i>(in corroboration).</i> Ugh, ugh, wah.</p>

<blockquote> <p><i>(A fire is lit and they dance round and over it till they seem part of the leaping flames.</i> TIGER LILY <i>invokes Manitou; the pipe of peace is broken; and they crawl off like a long snake that has not fed for many moons.</i> TOOTLES <i>peers after the tail and summons the other boys, who issue from their holes.)</i></p> </blockquote>

<p>TOOTLES. They are gone.</p>

<p>SLIGHTLY <i>(almost losing confidence in himself).</i> I do wish Peter was here.</p>

<p>FIRST TWIN. H'sh! What is that? <i>(He is gazing at the lagoon and shrinks back.)</i> It is wolves, and they are chasing Nibs!</p>

<blockquote><i>(The baying wolves are upon them quicker than any boy can scuttle down his tree.)</i></blockquote>

<p>NIBS <i>(falling among his comrades).</i> Save me, save me!</p>

<p>TOOTLES. What should we do?</p>

<p>SECOND TWIN. What would Peter do?</p>

<p>SLIGHTLY. Peter would look at them through his legs; let us do what Peter would do.</p>

<blockquote> <p><br> <i>(The boys advance backwards, looking between their legs at the snarling red-eyed enemy, who trot away foiled.)</i></p> </blockquote>

<p>FIRST TWIN <i>(swaggering).</i> We have saved you, Nibs. Did you see the pirates?</p>

<p>NIBS <i>(sitting up, and agreeably aware that the centre of interest is now to pass to him).</i> No, but I saw a wonderfuller thing, Twin. <i>(All mouths open for the information to be dropped into them.)</i> High over the lagoon I saw the loveliest great white bird. It is flying this way. <i>(They search the firmament.)</i></p>

<p>TOOTLES. What kind of a bird, do you think?</p>

<p>NIBS <i>(awed).</i> I don't know; but it looked so weary, and as it flies it moans 'Poor Wendy.'</p>

<p>SLIGHTLY <i>(instantly).</i> I remember now there are birds called Wendies.</p>

<p>FIRST TWIN <i>(who has flown to a high branch).</i> See, it comes, the Wendy! <i>(They all see it now.)</i> How white it is! <i>(A dot of light is pursuing the bird malignantly.)</i></p>

<p>TOOTLES. That is Tinker Bell. Tink is trying to hurt theWendy. <i>(He makes a cup of his hands and calls)</i> Hullo,Tink! <i>(A response comes down in the fairy language.)</i> She says Peter wants us to shoot the Wendy.</p>

<p>NIBS. Let us do what Peter wishes.</p>

<p>SLIGHTLY. Ay, shoot it; quick, bows and arrows.</p>

<p>TOOTLES <i>(first with his bow).</i> Out of the way, Tink; I'll shoot it. <i>(His bolt goes home, and</i> WENDY, <i>who has been fluttering among the tree-tops in her white nightgown, falls</i> <i>straight to earth. No one could be more proud than</i> TOOTLES.) I have shot the Wendy; Peter will be so pleased. <i>(From some tree on which</i> TINK <i>is roosting comes the tinkle we can now translate, 'You silly ass.'</i> TOOTLES <i>falters.)</i> Why do you say that? <i>(The others feel that he may have blundered, and draw away from</i> TOOTLES.)</p>

<p>SLIGHTLY <i>(examining the fallen one more minutely).</i> This is no bird; I think it must be a lady.</p>

<p>NIBS <i>(who would have preferred it to be a bird).</i> And Tootles has killed her.</p>

<p>CURLY. Now I see, Peter was bringing her to us. <i>(They wonder for what object.)</i></p>

<p>SECOND TWIN. To take care of us? <i>(Undoubtedly for some diverting purpose.)</i></p>

<p>OMNES <i>(though every one of them had wanted to have a shot at her).</i> Oh, Tootles!</p>

<p>TOOTLES <i>(gulping).</i> I did it. When ladies used to come to me in dreams I said 'Pretty mother,' but when she really came I shot her! <i>(He perceives the necessity of a solitary life for him.)</i> Friends, good-bye.</p>

<p>SEVERAL <i>(not very enthusiastic).</i> Don't go.</p>

<p>TOOTLES. I must; I am so afraid of Peter.</p>

<blockquote><i>(He has gone but a step toward oblivion when he is stopped by a crowing as of some victorious cock.)</i></blockquote>

<p>OMNES. Peter!</p>

<blockquote><i>(They make a paling of themselves in front of</i> WENDY <i>as</i> PETER <i>skims round the tree-tops and reaches earth.)</i></blockquote>

<p>PETER. Greeting, boys! <i>(Their silence chafes him.)</i> I am back; why do you not cheer? Great news, boys, I have brought at last a mother for us all.</p>

<p>SLIGHTLY <i>(vaguely).</i> Ay, ay.</p>

<p>PETER. She flew this way; have you not seen her?</p>

<p>SECOND TWIN <i>(as</i> PETER <i>evidently thinks her important).</i>Oh mournful day!</p>

<p>TOOTLES <i>(making a break in the paling).</i> Peter, I will show her to you.</p>

<p>THE OTHERS <i>(closing the gap).</i> No, no.</p>

<p>TOOTLES <i>(majestically).</i> Stand back all, and let Peter see.</p>

<blockquote><i>(The paling dissolves, and</i> PETER <i>sees</i> WENDY <i>prone on the ground.)</i></blockquote>

<p>PETER. Wendy, with an arrow in her heart! <i>(He plucks it out.)</i> Wendy is dead. <i>(He is not so much pained as puzzled.)</i></p>

<p>CURLY. I thought it was only flowers that die.</p>

<p>PETER. Perhaps she is frightened at being dead? <i>(Noneof them can say as to that.)</i> Whose arrow? <i>(Not one of them looks at</i> TOOTLES.)</p>

<p>TOOTLES. Mine, Peter.</p>

<p>PETER <i>(raising it as a dagger).</i> Oh dastard hand!</p>

<p>TOOTLES <i>(kneeling and baring his breast).</i> Strike, Peter; strike true.</p>

<p>PETER <i>(undergoing a singular experience).</i> I cannot strike; there is something stays my hand.</p>

<blockquote><i>(In fact</i> WENDY'S <i>arm has risen.)</i></blockquote>

<p>NIBS. 'Tis she, the Wendy lady. See, her arm. <i>(To help a friend)</i> I think she said 'Poor Tootles.'</p>

<p>PETER <i>(investigating).</i> She lives!</p>

<p>SLIGHTLY <i>(authoritatively).</i> The Wendy lady lives.<i>(The delightful feeling that they have been cleverer</i> <i>than they thought comes over them and they applaud</i> <i>themselves.)</i></p>

<p>PETER <i>(holding up a button that is attached to her chain).</i> See, the arrow struck against this. It is a kiss I gave her; it has saved her life.</p>

<p>SLIGHTLY. I remember kisses; let me see it. <i>(He takes it in his hand.)</i> Ay, that is a kiss.</p>

<p>PETER. Wendy, get better quickly and I'll take you to see the mermaids. She is awfully anxious to see a mermaid.</p>

<blockquote>(TINKER BELL, <i>who may have been off visiting her relations, returns to the wood and, under the impression that</i>WENDY <i>has been got rid of, is whistling as gaily as a canary. She is not wholly heartless, but is so small that she has only room for one feeling at a time.)</i></blockquote>

CURLY. Listen to Tink rejoicing because she thinks theWendy is dead! <i>(Regardless of spoiling another's pleasure)</i> Tink, the Wendy lives. <br> <blockquote>(TINK <i>gives expression to fury.)</i></blockquote>

<p>SECOND TWIN <i>(tell-tale).</i> It was she who said that you wanted us to shoot the Wendy.</p>

<p>PETER. She said that? Then listen, Tink, I am your friend no more. <i>(There is a note of acerbity in</i> TINK'S <i>reply; it may mean 'Who wants you?')</i> Begone from me forever. <i>(Now it is a very wet tinkle.)</i></p>

<p>CURLY. She is crying.</p>

<p>TOOTLES. She says she is your fairy.</p>

<p>PETER <i>(who knows they are not worth worrying about).</i> Oh well, not for ever, but for a whole week.</p>

<blockquote>(TINK. <i>goes off sulking, no doubt with the intention ofgiving all her friends an entirely false impression of</i>WENDY'S <i>appearance.)</i></blockquote>

<p>Now what shall we do with Wendy?</p>

<p>CURLY. Let us carry her down into the house.</p>

<p>SLIGHTLY. Ay, that is what one does with ladies.</p>

<p>PETER. No, you must not touch her; it wouldn't be sufficiently respectful.</p>

<p>SLIGHTLY. That is what I was thinking.</p>

<p>TOOTLES. But if she lies there she will die.</p>

<p>SLIGHTLY. Ay, she will die. It is a pity, but there is no way out.</p>

<p>PETER. Yes, there is. Let us build a house around her! <i>(Cheers again, meaning that no difficulty baffles</i> PETER.) Leave all to me. Bring the best of what we have. Gut our house. Be sharp. <i>(They race down their trees.)</i></p>

<blockquote><i>(While</i> PETER <i>is engrossed in measuring</i> WENDY <i>so that the house may fit her,</i> JOHN <i>and</i> MICHAEL, <i>who have probably landed on the island with a bump, wander forward, so draggled and tired that if you were to ask</i> MICHAEL <i>whether he is awake or asleep he would probably answer 'I haven't tried yet.')</i></blockquote>

<p>MICHAEL <i>(bewildered).</i> John, John, wake up. Where is Nana, John?</p>

<p>JOHN <i>(with the help of one eye but not always the same eye).</i> It is true, we did fly! <i>(Thankfully)</i> And here is Peter. Peter, is this the place?</p>

<blockquote>(PETER, <i>alas, has already forgotten them, as soon maybe he will forget</i> WENDY. <i>The first thing she should do now that she is here is to sew a handkerchief for him, and knot it as a jog to his memory.)</i></blockquote>

<p>PETER <i>(curtly).</i> Yes.</p>

<p>MICHAEL. Where is Wendy? (PETER <i>points.)</i></p>

<p>JOHN <i>(who still wears his hat).</i> She is asleep.</p>

<p>MICHAEL. John, let us wake her and get her to make supper for us.</p>

<blockquote><i>(Some of the boys emerge, and he pinches one.)</i></blockquote>

John, look at them! 

<p>PETER <i>(still house-building).</i> Curly, see that these boy shelp in the building of the house.</p>

<p>JOHN. Build a house?</p>

<p>CURLY. For the Wendy.</p>

<p>JOHN <i>(feeling that there must be some mistake here).</i> ForWendy? Why, she is only a girl.</p>

<p>CURLY. That is why we are her servants.</p>

<p>JOHN <i>(dazed).</i> Are you Wendy's servants?</p>

<p>PETER. Yes, and you also. Away with them. <i>(In another moment they are woodsmen hacking at trees, with</i> CURLY <i>as overseer.)</i> Slightly, fetch a doctor. (SLIGHTLY <i>reels and goes. He returns professionally in</i> JOHN'S <i>hat.)</i> Please, sir, are you a doctor?</p>

<p>SLIGHTLY <i>(trembling in his desire to give satisfaction).</i>Yes, my little man.</p>

<p>PETER. Please, sir, a lady lies very ill.</p>

<p>SLIGHTLY <i>(taking care not to fall over her).</i> Tut, tut, where does she lie?</p>

<p>PETER. In yonder glade. <i>(It is a variation of a game they play.)</i></p>

<p>SLIGHTLY. I will put a glass thing in her mouth. <i>(He inserts an imaginary thermometer in</i> WENDY'S <i>mouth and gives it a moment to record its verdict. He shakes it and then consults it.)</i></p>

<p>PETER <i>(anxiously).</i> How is she?</p>

<p>SLIGHTLY. Tut, tut, this has cured her.</p>

<p>PETER <i>(leaping joyously).</i> I am glad.</p>

<p>SLIGHTLY. I will call again in the evening. Give her beef tea out of a cup with a spout to it, tut, tut.</p>

<blockquote><i>(The boys are running up with odd articles of furniture.)</i></blockquote>

<p>PETER <i>(with an already fading recollection of the Darling nursery).</i> These are not good enough for Wendy. How Iwish I knew the kind of house she would prefer!</p>

<p>FIRST TWIN. Peter, she is moving in her sleep.</p>

<p>TOOTLES <i>(opening</i> WENDY'S <i>mouth and gazing down into the depths).</i> Lovely!</p>

<p>PETER. Oh, Wendy, if you could sing the kind of house you would like to have.</p>

<blockquote><i>(It is as if she had heard him.)</i></blockquote>

<p>WENDY <i>(without opening her eyes).</i></p>

<blockquote> I wish I had a woodland house,<br> The littlest ever seen,<br> With funny little red walls<br> And roof of mossy green.</blockquote>

<blockquote> <p><i>(In the time she sings this and two other verses, such is the urgency of</i> PETER'S <i>silent orders that they have knocked down trees, laid a foundation and put up the walls and roof, so that she is now hidden from view. 'Windows' cries</i> PETER, <i>and</i> CURLY <i>rushes them in, 'Roses' and</i> TOOTLES <i>arrives breathless with a festoon for the door. Thus springs into existence the most delicious little house for beginners.)</i></p> </blockquote>

<p>FIRST TWIN. I think it is finished.</p>

<p>PETER. There is no knocker on the door. (TOOTLES <i>hangs up the sole of his shoe.)</i> There is no chimney, we must have a chimney. <i>(They await his deliberations anxiously.)</i></p>

<p>JOHN <i>(unwisely critical).</i> It certainly does need a chimney.</p>

<blockquote><i>(He is again wearing his hat, which</i> PETER <i>seizes, knocks the top off it and places on the roof. In the friendliestway smoke begins to come out of the hat.)</i></blockquote>

<p>PETER <i>(with his hand on the knocker).</i> All look your best; the first impression is awfully important. <i>(he knocks, and after a dreadful moment of suspense, in which they cannot help wondering if any one is inside, the door opens and who should come out but</i> WENDY! <i>She has evidently been tidying a little. She is quite surprised to find that she has nine children.)</i></p>

<p>WENDY <i>(genteelly).</i> Where am I?</p>

<p>SLIGHTLY. Wendy lady, for you we built this house.</p>

<p>NIBS and TOOTLES. Oh, say you are pleased.</p>

<p>WENDY <i>(stroking the pretty thing).</i> Lovely, darling house!</p>

<p>FIRST TWIN. And we are your children.</p>

<p>WENDY <i>(affecting surprise).</i> Oh?</p>

<p>OMNES <i>(kneeling, with outstretched arms).</i> Wendy lady, be our mother! <i>(Now that they know it is pretend they acclaim her greedily.)</i></p>

<p>WENDY <i>(not to make herself too cheap).</i> Ought I? Of course it is frightfully fascinating; but you see I am only a little girl; I have no real experience.</p>

<p>OMNES. That doesn't matter. What we need is just a nice motherly person.</p>

<p>WENDY. Oh dear, I feel that is just exactly what I am.</p>

<p>OMNES. It is, it is, we saw it at once.</p>

<p>WENDY. Very well then, I will do my best. <i>(In their glee they go dancing obstreperously round the little house, and she sees she must be firm with them as well as kind.)</i> Come inside at once, you naughty children, I am sure your feet are damp. And before I put you to bed I have just time to finish the story of Cinderella.</p>

<blockquote> <p><i>(They all troop into the enchanting house, whose not least remarkable feature is that it holds them. A vision of</i> LIZA <i>passes, not perhaps because she has any right to be there; but she has so few pleasures and is so young that we just let her have a peep at the little house. By and by</i> PETER <i>comes out and marches up and down with drawn sword, for the pirates can be heard carousing faraway on the lagoon, and the wolves are on the prowl. The little house, its walls so red and its roof so mossy, looks very cosy</i> <i>and safe, with a bright light showing through the blind, the chimney smoking beautifully, and</i> PETER <i>on guard. On our last sight of him it is so dark that we just guess he is the little figure who has fallen asleep by the door. Dots of light come and go. They are inquisitive fairies having a look at the house. Any other child in their way they would mischief, but they just tweak</i> PETER'S <i>nose and pass on. Fairies, you see,can touch him.)</i></p> </blockquote>
