% !TEX program = pdflatex
% !TEX encoding = UTF-8
% !TEX spellcheck = en_GB
% !TEX root = peter-pan.tex

\Act{The Mermaids’ Lagoon}

\endinput

<p><i>It is the end of a long playful day on the lagoon. The sun's rays have persuaded him to give them another five minutes, for one more race over the waters before he gathers them up and lets in the moon. There are many mermaids here, going plop-plop, and one might attempt to count the tails did they not flash and disappear so quickly. At times a lovely girl leaps in the air seeking to get rid of her excess of scales, which fall in a silver shower as she shakes them off. From the coral grottoes beneath the lagoon, where are the mermaids' bedchambers, comes fitful music.</i></p>

<p><i>One of the most bewitching of these blue-eyed creatures is lying lazily on Marooners' Rock, combing her long tresses and noting effects in a transparent shell. Peter and his band are in the water unseen behind the rock, whither they have tracked her as if she were a trout, and at a signal ten pairs of arms come whack upon the mermaid to enclose her. Alas, this is only what was meant to happen, for she hears the signal (which is the crow of a cock) and slips through their arms into the water. It has been such a near thing that there are scales on some of their hands. They climb on to the rock crestfallen.</i></p>

<p>WENDY <i>(preserving her scales as carefully as if they were rare postage stamps).</i> I did so want to catch a mermaid.</p>

<p>PETER <i>(getting rid of his).</i> It is awfully difficult to catch a mermaid.</p>

<blockquote><i>(The mermaids at times find it just as difficult to catch him, though he sometimes joins them in their one game,which consists in lazily blowing their bubbles into the air and seeing who can catch them. The number of bubbles</i> PETER <i>has flown away with! When the weather grows</i> <i>cold mermaids migrate 'to the other side of the world, and he once went with a great shoal of them half the way.)</i></blockquote>

<p>They are such cruel creatures, Wendy, that they try to pull boys and girls like you into the water and drown them.</p>

<p>WENDY <i>(too guarded by this time to ask what he means 'precisely by 'like you,' though she is very desirous of knowing).</i> How hateful!</p>

<blockquote><i>(She is slightly different in appearance now, rather rounder, while</i> JOHN <i>and</i> MICHAEL <i>are not quite so round. The reason is that when new lost children arrive at his underground home</i> PETER <i>finds new trees for them to go up and down by, and instead of fitting the tree to them he makes them fit the tree. Sometimes it can be done by adding or removing garments, but if you are bumpy, or the tree is an odd shape, he has things done to you with a roller, and after that you fit.</i><br> 

<p><i>The other boys are now playing King of the Castle, throwing each other into the water, taking headers and so on; but these two continue to talk.)</i></p> </blockquote>

<p>PETER. Wendy, this is a fearfully important rock. It is called Marooners' Rock. Sailors are marooned, you know, when their captain leaves them on a rock and sails away.</p>

<p>WENDY. Leaves them on this little rock to drown?</p>

<p>PETER <i>(lightly).</i> Oh, they don't live long. Their hands are tied, so that they can't swim. When the tide is full this rock is covered with water, and then the sailor drowns.</p>

<blockquote>(WENDY <i>is uneasy as she surveys the rock, which is the only one in the lagoon and no larger than a table. Since she last looked around a threatening change has come over the scene. The sun has gone, but the moon has not come. What has come is a cold shiver across the waters which has sent all the wiser mermaids to their coral recesses. They know that evil is creeping over the lagoon. Of the boys</i> PETER <i>is of course the first to scent it, and he has leapt to his feet before the words strike the rock&mdash;</i></blockquote>

<blockquote> 'And if we 're parted by a shot<br> We 're sure to meet below.'</blockquote>

<blockquote><i>The games on the rock and around it end so abruptly that several divers are checked in the air. There they hang waiting for the word of command from</i> PETER.<i>When they get it they strike the water simultaneously, and the rock is at once as bare as if suddenly they had been blown off it. Thus the pirates find it deserted when their dinghy strikes the rock and is nearly stove in by the concussion.)</i></blockquote>

<p>SMEE. Luff, you spalpeen, luff! <i>(They are</i> SMEE <i>and</i> STARKEY, <i>with</i> TIGER LILY, <i>their captive, bound hand and foot.)</i> What we have got to do is to hoist the redskin on to the rock and leave her there to drown.</p>

<blockquote><i>(To one of her race this is an end darker than death by fire or torture, for it is written in the laws of the Piccaninnies that there is no path through water to the happy hunting ground. Yet her face is impassive; she is the daughter of a chief and must die as a chief's daughter; it is enough.)</i></blockquote>

<p>STARKEY <i>(chagrined because she does not mewl).</i> No mewling. This is your reward for prowling round the ship with a knife in your mouth.</p>

<p>TIGER LILLY <i>(stoically).</i> Enough said.</p>

<p>SMEE <i>(who would have preferred a farewell palaver).</i> So that's it! On to the rock with her, mate.</p>

<p>STARKEY <i>(experiencing for perhaps the last time the stirrings of a man).</i> Not so rough, Smee; roughish, but not so rough.</p>

<p>SMEE <i>(dragging her on to the rock).</i> It is the captain's orders.</p>

<blockquote><i>(A stave has in some past time been driven into the rock, probably to mark the burial place of hidden treasure, and to this they moor the dinghy.)</i></blockquote>

WENDY <i>(in the water).</i> Poor Tiger Lily! 

<p>STARKEY. What was that? <i>(The children bob.)</i></p>

<p>PETER <i>(who can imitate the captain's voice so perfectly that even the author has a dizzy feeling that at times he was really H</i>OOK). Ahoy there, you lubbers!</p>

<p>STARKEY. It is the captain; he must be swimming out to us.</p>

<p>SMEE <i>(calling).</i> We have put the redskin on the rock,Captain.</p>

<p>PETER. Set her free.</p>

<p>SMEE. But, Captain&mdash;&mdash;</p>

<p>PETER. Cut her bonds, or I 'll plunge my hook in you.</p>

<p>SMEE. This is queer:</p>

<p>STARKEY <i>(unmanned).</i> Let us follow the captain's orders.</p>

<blockquote><i>(They undo the thongs and</i> TIGER LILY <i>slides between their legs into the lagoon, forgetting in her haste to utter her war-cry, but</i> PETER <i>utters it for her, so naturally that even the lost boys are deceived. It is at this moment that the voice of the true</i> HOOK <i>is heard.)</i></blockquote>

<p>HOOK. Boat ahoy!</p>

<p>SMEE <i>(relieved).</i> It is the captain.</p>

<blockquote>(HOOK <i>is swimming, and they help him to scale the rock. He is in gloomy mood.)</i></blockquote>

<p>STARKEY. Captain, is all well?</p>

<p>SMEE. He sighs.</p>

<p>STARKEY. He sighs again.</p>

<p>SMEE <i>(counting).</i> And yet a third time he sighs. <i>(With foreboding)</i> What's up, Captain?</p>

<p>HOOK <i>(who has perhaps found the large rich damp cake untouched).</i> The game is up. Those boys have found a mother!</p>

<p>STARKEY. Oh evil day!</p>

<p>SMEE. What is a mother?</p>

<p>WENDY <i>(horrified).</i> He doesn't know!</p>

<p>HOOK <i>(sharply).</i> What was that?</p>

<blockquote>(PETER <i>makes the splash of a mermaid's tail.)</i></blockquote>

<p>STARKEY. One of them mermaids.</p>

<p>HOOK. Dost not know, Smee? A mother is&mdash;&mdash; <i>(He finds it more difficult to explain than he had expected, and looks about him for an illustration. He finds one in a great bird which drifts past in a nest as large as the roomiest basin)</i> There is a lesson in mothers for you! The nest must have fallen intothe water, but would the bird desert her eggs? (PETER, <i>who is now more or less off his head, makes the sound of a bird answering in the negative. The nest is borne out of sight.)</i></p>

<p>STARKEY. Maybe she is hanging about here to protect Peter?</p>

<blockquote>(HOOK'S <i>face clouds still further and</i> PETER <i>just manages not to call out that he needs no protection.)</i></blockquote>

<p>SMEE <i>(not usually a man of ideas).</i> Captain, could we not kidnap these boys' mother and make her our mother?</p>

<p>HOOK. Obesity and bunions, 'tis a princely scheme. We will seize the children, make them walk the plank, and Wendy shall be our mother!</p>

<p>WENDY. Never! <i>(Another splash from</i> PETER.)</p>

<p>HOOK. What say you, bullies?</p>

<p>SMEE. There is my hand on 't.</p>

<p>STARKEY. And mine.</p>

<p>HOOK. And there is my hook. Swear. <i>(All swear.)</i> But I had forgot; where is the redskin?</p>

<p>SMEE <i>(shaken).</i> That is all right, Captain; we let her go.</p>

<p>HOOK <i>(terrible).</i> Let her go?</p>

<p>SMEE. 'Twas your own orders, Captain.</p>

<p>STARKEY <i>(whimpering).</i> You called over the water to us to let her go.</p>

<p>HOOK. Brimstone and gall, what cozening is here? <i>(Disturbed by their faithful faces)</i> Lads, I gave no such order.</p>

<p>SMEE 'Tis passing queer.</p>

<p>HOOK <i>(addressing the immensities).</i> Spirit that haunts thisdark lagoon to-night, dost hear me?</p>

<p>PETER <i>(in the same voice).</i> Odds, bobs, hammer and tongs, I hear you.</p>

<p>HOOK <i>(gripping the stave for support).</i> Who are you, stranger, speak.</p>

<p>PETER <i>(who is only too ready to speak).</i> I am Jas Hook, Captain of the <i>Jolly Roger.</i></p>

<p>HOOK <i>(now white to the gills).</i> No, no, you are not.</p>

<p>PETER. Brimstone and gall, say that again and I 'll cast anchor in you.</p>

<p>HOOK. If you are Hook, come tell me, who am I?</p>

<p>PETER. A codfish, only a codfish.</p>

<p>HOOK <i>(aghast).</i> A codfish?</p>

<p>SMEE <i>(drawing back from him).</i> Have we been captained all this time by a codfish?</p>

<p>STARKEY. It's lowering to our pride.</p>

<p>HOOK <i>(feeling that his ego is slipping from him).</i> Don't desert me, bullies.</p>

<p>PETER <i>(top-heavy).</i> Paw, fish, paw!</p>

<blockquote><i>(There is a touch of the feminine in</i> HOOK, <i>as in all the greatest prates, and it prompts him to try the guessing game.)</i></blockquote>

<p>HOOK. Have you another name?</p>

<p>PETER <i>(falling to the lure).</i> Ay, ay.</p>

<p>HOOK <i>(thirstily).</i> Vegetable?</p>

<p>PETER. No.</p>

<p>HOOK. Mineral?</p>

<p>PETER. No.</p>

<p>HOOK. Animal?</p>

<p>PETER <i>(after a hurried consultation with</i> TOOTLES). Yes.</p>

<p>HOOK. Man?</p>

<p>PETER <i>(with scorn).</i> No.</p>

<p>HOOK. Boy?</p>

<p>PETER, Yes.</p>

<p>HOOK. Ordinary boy?</p>

<p>PETER. No!</p>

<p>HOOK. Wonderful boy?</p>

<p>PETER <i>(to</i> WENDY'S <i>distress).</i> Yes!</p>

<p>HOOK. Are you in England?</p>

<p>PETER. No.</p>

<p>HOOK. Are you here?</p>

<p>PETER. Yes.</p>

<p>HOOK <i>(beaten, though he feels he has very nearly got it).</i> Smee, you ask him some questions.</p>

<p>SMEE <i>(rummaging his brains).</i> I can't think of a thing,</p>

<p>PETER. Can't guess, can't guess! <i>(Foundering in his cockiness)</i> Do you give it up?</p>

<p>HOOK <i>(eagerly).</i> Yes.</p>

<p>PETER. All of you?</p>

<p>SMEE and STARKEY. Yes.</p>

<p>PETER <i>(crowing).</i> Well, then, I am Peter Pan!</p>

<blockquote><i>(Now they have him.)</i></blockquote>

<p>HOOK. Pan! Into the water, Smee. Starkey, mind the boat. Take him dead or alive!</p>

<p>PETER <i>(who still has all his baby teeth).</i> Boys, lam into the pirates!</p>

<blockquote>For <i>a moment the only two we can see are in the dinghy, where</i> JOHN <i>throws himself on</i> STARKEY. STARKEY <i>wriggles into the lagoon and</i> JOHN <i>leaps so quickly after him that he reaches it first. The impression left on</i> STARKEY <i>is that he is being attacked by the</i> TWINS. <i>The water becomes stained. The dinghy drifts away. Here and there a head shows in the water, and once it is the head of the crocodile. In the growing gloom some strike at their friends,</i> SLIGHTLY <i>getting</i> TOOTLES <i>in the fourth rib while he himself is pinked by</i> CURLY. <i>It looks as if the boys were getting the worse of it, which is perhaps just as well at this point, because</i> PETER, <i>who will be the determining factor in the end, has a perplexing way of changing sides if he is winning too easily.</i> HOOK'S <i>iron claw makes a circle of black water round him from which opponents flee like fishes. There is only one prepared to enter that dreadful circle. His name is</i> PAN. <i>Strangely, it is not in the water that they meet.</i> HOOK <i>has risen tothe rock to breathe, and at the same moment</i> PETER <i>scales it on the opposite side. The rock is now wet and as slippery as a ball, and they have to crawl rather than climb. Suddenly they are face to face.</i> PETER <i>gnashes his pretty teeth with joy, and is gathering himself for the spring when he sees he is higher up the rock than his foe. Courteously he waits;</i> HOOK <i>sees his intention, and taking advantage of it claws twice.</i> PETER <i>is untouched, but unfairness is what he never can get used to, and in his bewilderment he rolls off the rock. The crocodile, whose tick has been drowned in the strife, rears its jaws, and</i> HOOK, <i>who has almost stepped into them, is pursued by it to land. All is quiet on the lagoon now, not a sound save little waves nibbling at the rock, which is smaller than when we last looked at it. Two boys appear with the dinghy, and the others despite their wounds climb into it. They send the cry 'Peter&mdash;Wendy' across the waters, but no answer comes.)</i></blockquote>

<p>NIBS. They must be swimming home.</p>

<p>JOHN. Or flying.</p>

<p>FIRST TWIN. Yes, that is it. Let us be off and call to them as we go.</p>

<blockquote><i>(The dinghy disappears with its load, whose hearts would sink it if they knew of the peril of</i> WENDY <i>and her captain. From near and far away come the cries 'Peter&mdash;Wendy' till we no longer hear them.</i></blockquote>

<blockquote><i>Two small figures are now on the rock, but they have fainted. A mermaid who has dared to come back in the stillness stretches up her arms and is slowly pulling</i>WENDY <i>into the water to drown her.</i> WENDY <i>starts up just in time.)</i></blockquote>

<p>WENDY. Peter! <i>(He rouses himself and looks around him.)</i> Where are we, Peter?</p>

<p>PETER. We are on the rock, but it is getting smaller. Soon the water will be over it. Listen!</p>

<blockquote><i>(They can hear the wash of the relentless little waves.)</i></blockquote>

<p>WENDY. We must go.</p>

<p>PETER. Yes.</p>

<p>WENDY. Shall we swim or fly?</p>

<p>PETER. Wendy, do you think you could swim or fly to the island without me?</p>

<p>WENDY. You know I couldn't, Peter, I am just a beginner.</p>

<p>PETER. Hook wounded me twice. <i>(He believes it; he is so good at pretend that he feels the pain, his arms hang limp.)</i> I can neither swim nor fly.</p>

<p>WENDY. Do you mean we shall both be drowned?</p>

<p>PETER. Look how the water is rising!</p>

<blockquote><i>(They cover their faces with their hands. Something touches</i> WENDY <i>as lightly as a kiss.)</i></blockquote>

<p>PETER <i>(with little interest).</i> It must be the tail of the kite we made for Michael; you remember it tore itself outof his hands and floated away. <i>(He looks up and sees the kite sailing overhead.)</i> The kite! Why shouldn't it carry you? <i>(He grips the tail and pulls, and the kite responds.)</i></p>

<p>WENDY. Both of us!</p>

<p>PETER. It can't lift two, Michael and Curly tried.</p>

<blockquote><i>(She knows very well that if it can lift her it can lift him also, for she has been told by the boys as a deadly secret that one of the queer things about him is that he is no weight at all. But it is a forbidden subject.)</i></blockquote>

<p>WENDY. I won't go without you. Let us draw lots which is to stay behind.</p>

<p>PETER. And you a lady, never! <i>(The tail is in her hands, and the kite is tugging hard. She holds out her mouth to</i> PETER, <i>but he knows they cannot do that.)</i> Ready, Wendy!</p>

<blockquote><i>(The kite draws her out of sight across the lagoon.</i><br> <i> The waters are lapping over the rock now, and</i> PETER <i>knows that it will soon be submerged. Pale rays of light mingle with the moving clouds, and from the coral grottoes is to be heard a sound, at once the most musical and the most melancholy in the Never Land, the mermaids calling to the moon to rise.</i> PETER <i>is afraid at last, and a tremor runs through him, like a shudder passing over the lagoon; but on the lagoon one shudder follows another till there are hundreds of them, and he feels just the one.)</i></blockquote>

<p>PETER <i>(with a drum beating in his breast as if he were a real boy at last).</i> To die will be an awfully big adventure.</p>

<blockquote> <p><i>(The blind rises again, and the lagoon is now suffused with moonlight. He is on the rock still, but the water is over his feet. The nest is borne nearer, and the bird, after cooing a message to him, leaves it and wings her way upwards.</i> PETER, <i>who knows the bird language, slips into the nest, first removing the two eggs and placing them in</i> STARKEY'S <i>hat, which has been left on the stave.The hat drifts away from the rock, but he uses the stave as a mast. The wind is driving him toward the open sea. He takes off his shirt, which he had forgotten to remove while bathing, and unfurls it as a sail. His vessel tacks, and he passes from sight, naked and victorious. The bird returns and sits on the hat.)</i></p> </blockquote>
