% !TEX program = pdflatex
% !TEX encoding = UTF-8
% !TEX spellcheck = en_GB
% !TEX root = peter-pan.tex

\Act{The Home under the Ground}

\endinput

<p><i>We see simultaneously the home under the ground, with the
children in it and the wood above ground with the redskins on it.
Below, the children are gobbling their evening meal; above, the
redskins are squatting in their blankets near the little house
guarding the children from the pirates. The only way of communicating
between these two parties is by means of the hollow trees.</i></p>

<p><i>The home has an earthen floor, which is handy for digging in if
you want to go fishing; and owing to there being so many entrances
there is not much wall space. The table at which the lost ones are
sitting is a board on top of a live tree trunk, which has been cut
flat but has such growing pains that the board rises as they eat, and
they have sometimes to pause in their meals to cut a bit more off the
trunk. Their seats are pumpkins or the large gay mushrooms of which
we have seen an imitation one concealing the chimney. There is an
enormous fireplace which is in almost any part of the room where you
care to light it, and across this Wendy has stretched strings, made
of fibre, from which she hangs her washing. There are also various
tomfool things in the room of no use whatever.</i></p>

<p><i>Michaels basket bed is nailed high up on the wall as if to
protect him from the cat, but there is no indication at present of
where the others sleep. At the back between two of the tree trunks is
a grindstone, and near it is a lovely hole, the size of a band-box,
with a gay curtain drawn across so that you cannot see what is
inside. This is Tink's withdrawing-room and bed-chamber, and it is
just as well that you cannot see inside, for it is so exquisite in
its decoration and in the personal apparel spread out on the bed that
you could scarcely resist making off with something. Tink is within
at present, as one can guess from a glow showing through the chinks.
It is her own glow, for though she has a chandelier for the look of
the thing, of course she lights her residence herself. She is
probably wasting valuable time just now wondering whether to put on
the smoky blue or the apple-blossom.</i></p>

<p><i>All the boys except Peter are here, and Wendy has the head of
the table, smiling complacently at their captivating ways, but doing
her best at the same time to see that they keep the rules about
hands-off-the-table, no-two-to-speak-at-once, and so on. She is
wearing romantic woodland garments, sewn by herself, with red berries
in her hair which go charmingly with her complexion, as she knows;
indeed she searched for red berries the morning after she reached the
island. The boys are in picturesque attire of her contrivance, and if
these don't always fit well the fault is not hers but the wearers,
for they constantly put on each other's things when they put on
anything at all. Michael is in his cradle on the wall. First Twin is
apart on a high stool and wears a dunce's cap, another invention of
Wendy's, but not wholly successful because everybody wants to be
dunce.</i></p>

<p><i>It is a pretend meal this evening, with nothing whatever on the
table, not a mug, nor a crust, nor a spoon. They often have these
suppers and like them on occasions as well as the other kind, which
consist chiefly of bread-fruit, tappa rolls, yams, mammee apples and
banana splash, washed down with calabashes of poe-poe. The pretend
meals are not Wendy's idea; indeed she was rather startled to find,
on arriving, that Peter knew of no other kind, and she is not
absolutely certain even now that he does eat the other kind, though
no one appears to do it more heartily. He insists that the pretend
meals should be partaken of with gusto, and we see his band doing
their best to obey orders.</i></p>

<p>WENDY <i>(her fingers to her ears, for their chatter and clatter
are deafening).</i> Silence! Is your mug empty, Slightly?</p>

<p>SLIGHTLY <i>(who would not say this if he had a mug).</i> Not
quite empty, thank you.</p>

<p>NIBS. Mummy, he has not even begun to drink his poe-poe.</p>

<p>SLIGHTLY <i>(seizing his chance, for this is tale-bearing).</i> I
complain of Nibs!<br>
</p>

<blockquote>(JOHN <i>holds up his hand.)</i></blockquote>

<p>WENDY. Well, John?</p>

<p>JOHN. May I sit in Peter's chair as he is not here?</p>

<p>WENDY. In your father's chair? Certainly not.</p>

<p>JOHN. He is not really our father. He did not even know how to be
a father till I showed him.<br>
</p>

<blockquote><i>(This is insurbordination.)</i></blockquote>

<p>SECOND TWIN. I complain of John!<br>
</p>

<blockquote><i>(The gentle</i> TOOTLED <i>raises his
hand.)</i></blockquote>

<p>TOOTLES <i>(who has the poorest opinion of himself).</i> I
don'tsuppose Michael would let me be baby?</p>

<p>MICHAEL. No, I won't.</p>

<p>TOOTLES. May I be dunce?</p>

<p>FIRST TWIN <i>(from his perch).</i> No. It's awfully difficultto
be dunce.</p>

<p>TOOTLES. As I can't be anything important would any of you like to
see me do a trick?</p>

<p>OMNES. No.</p>

<p>TOOTLES <i>(subsiding).</i> I hadn't really any hope.<br>
</p>

<blockquote><i>(The tale-telling breaks out again.)</i></blockquote>

<p>NIBS. Slightly is coughing on the table.</p>

<p>CURLY. The twins began wiih tappa rolls.</p>

<p>SLIGHTLY. I complain of Nibs!</p>

<p>NIBS. I complain of Slightly!</p>

<p>WENDY. Oh dear, I am sure I sometimes think that spinsters are to
be envied.</p>

<p>MICHAEL. Wendy, I am too big for a cradle.</p>

<p>WENDY. You are the littlest, and a cradle is such a nice homely
thing to have about a house. You others can clear away now. <i>(She
sits down on a pumpkin near the fire to her usual evening occupation,
darning.)</i> Every heel with a hole in it!<br>
</p>

<blockquote><i>(The boys clear away with dispatch, washing dishes
they don't have in a non-existent sink and stowing them ina cupboard
that isn't there. Instead of sawing the table-leg to-night they crush
it into the ground like a concertina, and are now ready for play, in
which they indulge hilariously.</i></blockquote>

<blockquote><i>A movement of the Indians draws our attention to the
scene above. Hitherto, with the exception of</i> PANTHER, <i>who sits
on guard on top of the little house, they have been hunkering in
their blankets, mute but picturesque; now all rise and prostrate
themselves before the majestic figure of</i> PETER, <i>who approaches
through the forest carrying a gun and game bag. It is not exactly a
gun. He often wanders away alone with this weapon, and when he comes
back you are never absolutely certain whether he has had an adventure
or not. He may have forgotten it so completely that he says nothing
about it; and then when you go out you find the body. On the other
hand he may say a great deal about it, and yet you never find the
body. Sometimes he comes home with his face scratched, and tells</i>
WENDY, <i>as a thing of no importance, that he got these marks from
the little people for cheeking them at a fairy wedding, and she
listens politely, but she is never quite sure, you know; indeed
theonly one who is sure about anything on the island is</i>
PETER.)</blockquote>

<p>PETER. The Great White Father is glad to see the Piccaninny braves
protecting his wigwam from the pirates.</p>

<p>TIGER LILY. The Great White Father save me from pirates. Me his
velly nice friend now; no let pirates hurt him.</p>

<p>BRAVES. Ugh, ugh, wah!</p>

<p>TIGER LILY. Tiger Lily has spoken.</p>

<p>PANTHER. Loola, loola! Great Big Little Panther has spoken.</p>

<p>PETER. It is well. The Great White Father has spoken.<br>
</p>

<blockquote><i>(This has a note of finality about it, with the
implied ,'And now shut up' which is never far from the courteous
receptions of well-meaning inferiors by born leaders of men. He
descends his tree, not unheard by</i> WENDY.)</blockquote>

<p>WENDY. Children, I hear your father's step. He likes you to meet
him at the door. (PETER <i>scatters pretend nuts among them and
watches sharply to see that they crunch with relish.)</i> Peter, you
just spoil them, you know!</p>

<p>JOHN <i>(who would be incredulous if he dare).</i> Any sport,
Peter?</p>

<p>PETER. Two tigers and a pirate.</p>

<p>JOHN <i>(boldly).</i> Where are their heads?</p>

<p>PETER <i>(contracting his little brows.)</i> In the bag.</p>

<p>JOHN. <i>(No, he doesn't say it. He backs away.)</i></p>

<p>WENDY <i>(peeping into the bag).</i> They are beauties'. <i>(She
has learned her lesson.)</i></p>

<p>FIRST TWIN. Mummy, we all want to dance.</p>

<p>WENDY. The mother of such an armful dance!</p>

<p>SLIGHTLY. As it is Saturday night?<br>
</p>

<blockquote><i>(They have long lost count of the days, but always if
they want to do anything special they say this is Saturday night, and
then they do it.)</i></blockquote>

<p>WENDY. Of course it is Saturday night, Peter? <i>(He shrugs an
indifferent assent.)</i> On with your nighties first.<br>
</p>

<blockquote><i>(They disappear into various recesses, and</i> PETER
<i>and</i> WENDY <i>with her darning are left by the fire to dodder
parentally. She emphasises it by humming a verse of 'John Anderson my
Jo,' which has not the desired effect on</i> PETER. <i>She is too
loving to be ignorant that he is not loving enough, and she hesitates
like one who knows the answer to her question.)</i></blockquote>

<p>What is wrong, Peter?</p>

<p>PETER <i>(scared).</i> It is only pretend, isn't it, that I am
their father?</p>

<p>WENDY <i>(drooling).</i> Oh yes.<br>
</p>

<blockquote><i>(His sigh of relief is without consideration for her
feelings.)</i></blockquote>

<p>But they are ours, Peter, yours and mine.</p>

<p>PETER <i>(determined to get at facts, the only things that puzzle
him).</i> But not really?</p>

<p>WENDY. Not if you don't wish it.</p>

<p>PETER. I don't.</p>

<p>WENDY <i>(knowing she ought not to 'probe but driven to it by
something within.)</i> What are your exact feelings for me,
Peter?</p>

<p>PETER <i>(in the class-room).</i> Those of a devoted
son,Wendy.</p>

<p>WENDY <i>(turning away).</i> I thought so.</p>

<p>PETER. You are so puzzling. Tiger Lily is just the same; there is
something or other she wants to be to me, but she says it is not my
mother.</p>

<p>WENDY <i>(with spirit).</i> No, indeed it isn't.</p>

<p>PETER. Then what is it?</p>

<p>WENDY. It isn't for a lady to tell.<br>
</p>

<blockquote><i>(The</i> <i>curtain of the fairy chamber opens
slightly, and</i> TINK, <i>who has doubtless been eavesdroping,
tinkles a laugh of scorn.)</i></blockquote>

<p>PETER <i>(badgered).</i> I suppose she means that she wants to be
my mother.<br>
</p>

<blockquote>(TINK'S <i>comment is 'You silly ass.')</i></blockquote>

<p>WENDY <i>(who has picked up some of the fairy words).</i> I almost
agree with her!<br>
</p>

<blockquote><i>(The arrival of the boys in their nightgowns turns</i>
WENDY'S <i>mind to practical matters, for the children have to be
arranged in line and passed or not passed for cleanliness.</i>
SLIGHTLY <i>is the worst. At last we see how they sleep, for in a
babel the great bed which stands on end by day against the wall is
unloosed from custody and lowered to the floor. Though large, it is a
tight fit for so many boys, and</i> WENDY <i>has made a rule that
there is to be no turning round until one gives the signal, when all
turn at once.</i></blockquote>

<blockquote>
<p>FIRST TWIN <i>is the best dancer and performs mightily on the bed
and in it and out of it and over it to an accompaniment of pillow
fights by the less agile; and then there is a rush at</i> WENDY.)</p>
</blockquote>

<p>NIBS. Now the story you promised to tell us as soon as we were in
bed!</p>

<p>WENDY <i>(severely).</i> As far as I can see you are not in bed
yet.<br>
</p>

<blockquote><i>(They scramble into the bed, and the effect is as of a
boxful of sardines.)</i></blockquote>

<p>WENDY <i>(drawing up her stool).</i> Well, there was once a
gentleman&mdash;&mdash;&mdash;</p>

<p>CURLY. I wish he had been a lady.</p>

<p>NIBS. I wish he had been a white rat.</p>

<p>WENDY. Quiet! There was a lady also. The gentleman's name was Mr.
Darling and the lady's name was Mrs. Darling&mdash;&mdash;&mdash;</p>

<p>JOHN. I knew them!</p>

<p>MICHAEL <i>(who has been allowed to join the circle).</i> I think
I knew them.</p>

<p>WENDY. They were married, you know; and what do you think they
had?</p>

<p>NIBS. White rats?</p>

<p>WENDY. No, they had three descendants. White rats are descendants
also. Almost everything is a descendant. Now these three children had
a faithful nurse called Nana.</p>

<p>MICHAEL <i>(alas).</i> What a funny name!</p>

<p>WENDY. But Mr. Darling&mdash;<i>(faltering)</i> or was it
Mrs.Darling?&mdash;was angry with her and chained her up in the yard;
so all the children flew away. They flew away to the Never Land,
where the lost boys are.</p>

<p>CURLY. I just thought they did; I don't know how it is, but I just
thought they did.</p>

<p>TOOTLES. Oh, Wendy, was one of the lost boys called Tootles.</p>

<p>WENDY. Yes, he was.</p>

<p>TOOTLES <i>(dazzled).</i> Am I in a story? Nibs, I am in a
story!</p>

<p>PETER <i>(who is by the fire making</i> Pan's <i>pipes with his
knife, and is determined that</i> WENDY <i>shall have fair pl</i>ay,
<i>however beastly a story he may think it).</i> A little less noise
there.</p>

<p>WENDY <i>(melting over the beauty of her present performance, but
without any real qualms).</i> Now I want you to consider the feelings
of the unhappy parents with all their children flown away. Think, oh
think, of the empty beds. <i>(The heartless ones think of them with
glee.)</i></p>

<p>FIRST TWIN <i>(cheerfully).</i> It's awfully sad.</p>

<p>WENDY. But our heroine knew that her mother would always leave the
window open for her progeny to fly back by; so they stayed away for
years and had a lovely time.<br>
</p>

<blockquote>(PETER <i>is</i> <i>interested at last.)</i></blockquote>

<p>FIRST TWIN. Did they ever go back?</p>

<p>WENDY <i>(comfortably).</i> Let us now take a peep into the
future. Years have rolled by, and who is this elegant lady of
uncertain age alighting at London station?<br>
</p>

<blockquote><i>(The tension is unbearable.)</i></blockquote>

<p>NIBS. Oh, Wendy, who is she?</p>

<p>WENDY <i>(swelling).</i> Can it be&mdash;yes&mdash;no&mdash;yes,
it is the fair Wendy!</p>

<p>TOOTLES. I am glad.</p>

<p>WENDY. Who are the two noble portly figures accompanying her? Can
they be John and Michael? They are. <i>(Pride of</i> MICHAEL.) 'See,
dear brothers,' says Wendy, pointing upward, 'there is the window
standing open.' So up they flew to their loving parents, and pen
cannot inscribe the happy scene over which we draw a veil. <i>(Her
triumph is spoilt by a groan from</i> PETER <i>and she hurries to
him.)</i> Peter, what is it? <i>(Thinking he is ill, and looking
lower than his chest.)</i> Where is it?</p>

<p>PETER. It isn't that kind of pain. Wendy, you are wrong about
mothers. I thought like you about the window, so I stayed away for
moons and moons, and then I flew back, but the window was barred, for
my mother had forgotten all about me and there was another little boy
sleeping in my bed.<br>
</p>

<blockquote>( <i>This is a general damper.)</i></blockquote>

<p>JOHN. Wendy, let us go back!</p>

<p>WENDY. Are you sure mothers are like that?</p>

<p>PETER. Yes.</p>

<p>WENDY. John, Michael! <i>(She clasps them to her.)</i></p>

<p>FIRST TWIN <i>(alarmed).</i> You are not to leave us, Wendy?</p>

<p>WENDY. I must.</p>

<p>NIBS. Not to-night?</p>

<p>WENDY. At once. Perhaps mother is in half-mourning by this time!
Peter, will you make the necessary arrangements?<br>
</p>

<blockquote><i>(She asks it in the steely tones women adopt when they
are prepared secretly for opposition.)</i></blockquote>

<p>PETER <i>(coolly).</i> If you wish it.<br>
</p>

<blockquote><i>(He ascends his tree to give the redskins their
instructions. The lost boys gather threateningly round</i>
WENDY.)</blockquote>

<p>CURLY. We won't let you go!</p>

<p>WENDY <i>(with one of those inspirations women have, in an
emergency, to make use of some male who need otherwise have no
hope).</i> Tootles, I appeal to you.</p>

<p>TOOTLES <i>(leaping to his death if necessary).</i> I am just
Tootles and nobody minds me, but the first who does not behave to
Wendy I will blood him severely. (PETER <i>returns.)</i></p>

<p>PETER <i>(with awful serenity).</i> Wendy, I told the braves to
guide you through the wood as flying tires you so. Then Tinker Bell
will take you across the sea. <i>(A shrill tinkle from the boudoir
probably means 'and drop her into it.')</i></p>

<p>NIBS <i>(fingering the curtain which he is not allowed to
open).</i> Tink, you are to get up and take Wendy on a journey.
<i>(Star-eyed)</i> She says she won't!</p>

<p>PETER <i>(taking a step toward that chamber).</i> If you don'tget
up, Tink, and dress at once&mdash;&mdash; She is getting up!</p>

<p>WENDY <i>(quivering now that the time to depart has come).</i>
Dear ones, if you will all come with me I feel almost sure I can get
my father and mother to adopt you.<br>
</p>

<blockquote><i>(There is joy at this, not that they want parents, but
novelty is their religion.)</i></blockquote>

<p>NIBS. But won't they think us rather a handful?</p>

<p>WENDY <i>(a swift reckoner).</i> Oh no, it will only mean having a
few beds in the drawing-room; they can be hidden behind screens on
first Thursdays.<br>
</p>

<blockquote><i>(Everything depends on</i> PETER.)</blockquote>

<p>OMNES. Peter, may we go?</p>

<p>PETER <i>(carelessly through the pipes to which he is giving a
finishing touch).</i> All right.<br>
</p>

<blockquote><i>(They scurry off to dress for the
adventure.)</i></blockquote>

<p>WENDY <i>(insinuatingly).</i> Get your clothes, Peter.</p>

<p>PETER <i>(skipping about and playing fairy music on his pipes, the
only music he knows).</i> I am not going with you,Wendy.</p>

<p>WENDY. Yes, Peter!</p>

<p>PETER. No.<br>
</p>

<blockquote><i>(The lost ones run back gaily, each carrying a stick
witha bundle on the end of it.)</i></blockquote>

<p>WENDY. Peter isn't coming!<br>
</p>

<blockquote><i>(All the faces go blank.)</i></blockquote>

<p>JOHN <i>(even</i> JOHN). Peter not coming!</p>

<p>TOOTLES <i>(overthrown).</i> Why, Peter?</p>

<p>PETER <i>(his pipes more riotous than ever).</i> I just want
always to be a little boy and to have fun.<br>
</p>

<blockquote><i>(There is a general fear that they are perhaps making
the mistake of their lives.)</i></blockquote>

<p>Now then, no fuss, no blubbering. <i>(With dreadful cynicism)</i>
I hope you will like your mothers! Are you ready, Tink? Then lead the
way.<br>
</p>

<blockquote>(TINK <i>darts up any tree, but she is the only one.
Theair above is suddenly rent with shrieks and the clash of steel.
Though they cannot see, the boys know that</i> HOOK<i>and his crew
are upon the Indians. Mouths open andremain open, all in mute appeal
to</i> PETER. <i>He is theonly boy on his feet now, a sword in his
hand, the samehe slew Barbicue with; and in his eye is the lust of
battle.</i></blockquote>

<blockquote><i>We can watch the carnage that is invisible to the
children.</i> HOOK <i>has basely broken the two laws of Indian
warfare, which are that the redskins should attack first, and that it
should be at dawn. They have known the pirate whereabouts since,
early in the night, one of</i> SMEE'S <i>fingers crackled. The
brushwood has closed behind their scouts as silently as the sand on
the mole; for hours they have imitated the lonely call of the coyote;
no stratagem has been overlooked, but alas, they have trusted to the
pale-face's honour to await an attack at dawn, when his courage is
known to be at the lowest ebb.</i> HOOK <i>falls upon them pell-mell,
and one cannot withhold a reluctani admiration for the wit that
conceived so subtle a scheme and the fell genius with which it is
carried out. If the braves would rise quickly they might still have
time to scalp, but this they are forbidden to do by the traditions of
their race, for it is written that they must never express surprise
in the presence of the pale-face. For a brief space they remain
recumbent, not a muscle moving, as if the foe were here by
invitation. Thus perish the flower of the Piccaninnies, though not
unavenged, for with</i> LEAN WOLF <i>fall</i> ALF MASON <i>and</i>
CANARY ROBB, <i>while other pirates to bite dust are</i> BLACK
GILMOUR <i>and</i> ALAN HERB, <i>that same</i> HERB <i>who is still
remembered at Manaos for playing skittles with the mate of the</i>
Switch <i>for each other's heads.</i> CHAY TURLEY, <i>who laughed
with the wrong side of his mouth (having no other), is tomahawked
by</i> PANTHER, <i>who eventually cuts a way through the shambles
with</i> TIGER LILY <i>and a remnant of the tribe.</i></blockquote>

<blockquote><i>This onslaught passes and is gone like a fierce wind.
The victors wipe their cutlasses, and squint, ferret-eyed, at their
leader. He remains, as ever, aloof in spirit and in substance. He
signs to them to descend the trees, for he is convinced that</i> PAN
<i>is down there, and though he has smoked the bees it is the honey
he wants. There is something in</i> PETER <i>that at all times goads
this extraordinary man to frenzy; it is the boy's cockiness, which
disturbs</i> HOOK <i>like an insect. If you have seen a lion in a
cage futilely pursuing a sparrow you will know what is meant. The
pirates try to do their captain's bidding, but the apertures prove to
be not wide enough for them; he cannot even ram them down with a
pole. He steals to the mouth of a tree and listens.)</i></blockquote>

<p>PETER <i>(prematurely).</i> All is over!</p>

<p>WENDY. But who has won?</p>

<p>PETER. Hst! If the Indians have won they will beat the tom-tom; it
is always their signal of victory.<br>
</p>

<blockquote>(HOOK <i>licks his lips at this and signs to</i> SMEE,
<i>who is sitting on it, to hold up the tom-tom. He beats upon it</i>
<i>with his claw, and listens for results.)</i></blockquote>

<p>TOOTLES. The tom-tom!</p>

<p>PETER <i>(sheathing his sword).</i> An Indian victory!<br>
</p>

<blockquote><i>(The cheers from below are music to the black hearts
above.)</i></blockquote>

<p>You are quite safe now, Wendy. Boys, good-bye. <i>(He resumes his
pipes.)</i></p>

<p>WENDY. Peter, you will remember about changing your flannels,
won't you?'</p>

<p>PETER. Oh, all right!</p>

<p>WENDY. And this is your medicine.<br>
</p>

<blockquote><i>(She puts something into a shell and leaves it on a
ledge between two of the trees. It is only water, but she measures it
out in drops.)</i></blockquote>

<p>PETER. I won't forget.</p>

<p>WENDY. Peter, what are you to me?</p>

<p>PETER <i>(through the pipes).</i> Your son, Wendy.</p>

<p>WENDY. Oh, good-bye!<br>
</p>

<blockquote><i>(The travellers start upon their journey, little
witting that</i> HOOK <i>has issued his silent orders: a man to the
mouth of each tree, and a row of men between the trees and the little
house. As the children squeeze up they are plucked from their trees,
trussed, thrown like bales of cotton from one pirate to another, and
so piled up in the little house. The only one treated differently
is</i> WENDY, <i>whom</i> HOOK <i>escorts to the house on his arm
with hateful politeness. He signs to his dogs to be gone, and they
depart through the wood, carrying the little house with its strange
merchandise and singing their ribald song. The chimney of the little
house emits a jet of smoke fitfully, as if not sure what it ought to
do just now.</i></blockquote>

<blockquote>HOOK <i>and</i> PETER <i>are now, as it were, alone on
the island. Below,</i> PETER <i>is on the bed, asleep, no weapon near
him; above,</i> HOOK, <i>armed to the teeth, is searching noiselessly
for some tree down which the nastiness of him can descend. Don't be
too much alarmed by this; it is precisely the situation</i> PETER
<i>would have chosen; indeed if the whole thing were
pretend&mdash;&mdash;. One of his arms droops over the edge of the
bed, a leg is arched, and the mouth is not so tightly closed that we
cannot see the little pearls. He is dreaming, and in his dreams he is
always in pursuit of a boy who was never here, nor anywhere: the only
boy who could beat him.</i></blockquote>

<blockquote>HOOK <i>finds the tree. It is the one set apart for</i>
SLIGHTLY <i>who being addicted when hot to the drinking of water has
swelled in consequence and surreptitiously scooped his tree for
easier descent and egress. Down this the pirate wriggles a passage.
In the aperture below his face emerges and goes green as he glares at
the sleeping child. Does no feeling of compassion disturb his sombre
breast? The man is not wholly evil: he has a</i> Thesaurus <i>in his
cabin, and is no mean performer on the flute. What really warps him
is a presentiment that he is about to fail. This is not unconnected
with a beatific smile on the face of the sleeper, whom he cannot
reach owing to being stuck at the foot of the tree. He, however, sees
the medicine shell within easy reach, and to</i> WENDY'S <i>draught
he adds from a bottle five drops of poison distilled when he was
weeping from the red in his eye. The expression on</i> PETER'S
<i>face merely implies that something heavenly is going on.</i> HOOK
<i>worms his way upwards, and winding his cloak around him, as if to
conceal his person from the night of which he is the blackest part,
he stalks moodily toward the lagoon.</i></blockquote>

<blockquote><i>A dot of light flashes past him and darts down the
nearest tree, looking for</i> PETER, <i>only for</i> PETER, <i>quite
indifferent about the others when she finds him
safe.)</i></blockquote>

<p>PETER <i>(stirring).</i> Who is that? (TINK <i>has to tell her
tale, in one long ungrammatical sentence.)</i> The redskins were
defeated? Wendy and the boys captured by the pirates! I'll rescue
her, I'll rescue her! <i>(He leap first at his dagger, and then at
his grindstone, to sharpen it.</i> TINK <i>alights near the shell,
and rings out a warning cry.)</i> Oh, that is just my medicine.
Poisoned? Who could have poisoned it? I promised Wendy to take it,
and I will as soon as I have sharpened my dagger. (TINK, <i>who sees
its red colour and remembers the red in the grate's eye, nobly
swallows the draught as</i> PETER'S <i>hand is reaching for it.)</i>
Why, Tink, you have drunk my medicine! <i>(She flutters strangely
about the room, answering him now in a very thin tinkle.)</i> It was
poisoned and you drank it to save my life! Tink, dear Tink, are you
dying? <i>(He has never called her dear</i> TINK <i>before, and for a
moment she is gay; she alights on his shoulder, gives his chin a
loving bite, whispers 'You silly ass' and falls on her tiny bed. The
boudoir, which is lit by her, flickers ominously. He is on his knees
by the opening.)</i></p>

<p>Her light is growing faint, and if it goes out, that means she is
dead! Her voice is so low I can scarcely tell what she is saying. She
says&mdash;she says she thinks she could get well again if children
believed in fairies! <i>(He rises and throws out his arms he knows
not to whom, perhaps to the boys and girls of whom he is not
one.)</i> Do you believe in fairies? Say quick that you believe! If
you believe, clap your hands! <i>(Many clap, some don't, a few hiss.
Then perhaps there is a rush of Nanas to the nurseries to see what on
earth is happening. But</i> TINK <i>is saved.)</i> Oh, thank you,
thank you, thank you! And now to rescue Wendy!<br>
</p>

<blockquote>
<p>(TINK <i>is already as merry and impudent as a grig, withnot a
thought for those who have saved her.</i> PETER <i>ascends his tree
as if he were shot up it. What he is feeling is</i> 'HOOK <i>or me
this time!' He is frightfully happy. He soon hits the trail, for the
smoke from the house has lingered here and there to guide him. He
takes wing.)</i></p>

<p> </p>
</blockquote>
